---
title: "Proteomic Analysis of the Glioblastoma Cell Hierarchy"
author: "Alejandro Pérez Recio"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing and loading the necessary libraries:

```{r}
# Install missing packages

cran_packages <- c(
  "readxl", "FactoMineR", "factoextra", "tidyverse",
  "dplyr", "tibble", "ggplot2", "ggrepel",
  "fgsea","pheatmap"
)

bioc_packages <- c(
  "NormalyzerDE", "limma", "clusterProfiler",
  "org.Hs.eg.db", "enrichplot", "GSVA",
  "msigdbr", "GSEABase", "decoupleR",
  "dorothea", "EnhancedVolcano", "mygene"
)

# Install BiocManager if needed
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}

# Install missing CRAN packages
missing_cran <- cran_packages[!cran_packages %in% installed.packages()[, "Package"]]
if (length(missing_cran) > 0) {
  install.packages(missing_cran)
}

# Install missing Bioconductor packages
missing_bioc <- bioc_packages[!bioc_packages %in% installed.packages()[, "Package"]]
if (length(missing_bioc) > 0) {
  BiocManager::install(missing_bioc, ask = FALSE, update = FALSE)
}

# Load packages

all_packages <- c(cran_packages, bioc_packages)

invisible(lapply(all_packages, library, character.only = TRUE))
```

Loading the dataset:

```{r}
prots_df <- read_excel("20260413_112142_Glioblastoma y RAD52_Report.xlsx",
                       na = c("", "NaN", "NA"))
head(prots_df)
```

Protein groups are aggregated at the gene level by averaging intensities for proteins mapping to the same gene, to enable compatibility with gene set–based pathway analysis.

We will also separate the samples into two different dataframes to normalyze them separately (stem-like and differentiated) because of the big proteomic differences between these two cell lines, so that we can compare the RAD52 samples separately. We will keep a joint dataframe for the GSC vs DGC comparison.

We will filter out those proteins with a high proportion of NA values. However, we should be careful with the filter, since NA values often derive from low protein quantity, and might be biologically relevant (for example, if the NA values for one protein are mostly in one condition, it is most probably due to that condition having low quantity of that protein, which is informative and relevant, and should not be simply discarded).

Next, we will impute data using Limit of Detection (LoD), using 1/5 of the minimum value per protein.

We use a boxplot to represent the global distribution of gene expression in the different samples and check for the presence of noise.

```{r}
# Since we will be working with the PG.Genes IDs, and some proteins in the DF
# share the same PG.Gene (they are isoforms, splice variants, etc.) we will
# collapse the proteins belonging to the same gene into one, using the mean
# expression

df_collapsed <- prots_df %>%
  dplyr::group_by(PG.Genes) %>%
  dplyr::summarise(across(where(is.numeric), mean))

protein_ids <- df_collapsed$PG.Genes

# The original DF contains other data and the replicates are not in adjacent columns.
prots_df_ordered <- data.frame(df_collapsed[2],df_collapsed[6],df_collapsed[10],df_collapsed[14],
                               df_collapsed[3],df_collapsed[7],df_collapsed[11],df_collapsed[15],
                               df_collapsed[4],df_collapsed[8],df_collapsed[12],df_collapsed[16],
                               df_collapsed[5],df_collapsed[9],df_collapsed[13],df_collapsed[17]
                               )

# One of the proteins (Q6ZSR9) has no PG.Gene associated with it (It is encoded
# in the 3'-UTR of AAK1 but there is evidence for the existence of the protein
# from a number of proteomics studies). We will remove this protein for this
# study as we will not be able to map it to a gene. It is located as the last
# row of the data.

prots_df_ordered <- prots_df_ordered[-nrow(prots_df_ordered), ]
protein_ids <- protein_ids[-length(protein_ids)]

rownames(prots_df_ordered) <- protein_ids
colnames(prots_df_ordered) <- c('shNT_DIFF_1','shNT_DIFF_2','shNT_DIFF_3','shNT_DIFF_4',
                                'shNT_STEM_1','shNT_STEM_2','shNT_STEM_3','shNT_STEM_4',
                                'shRAD52_DIFF_1','shRAD52_DIFF_2','shRAD52_DIFF_3','shRAD52_DIFF_4',
                                'shRAD52_STEM_1','shRAD52_STEM_2','shRAD52_STEM_3','shRAD52_STEM_4')

# Separate into 2 different dataframes for the two cell tzpes
prots_DIFF <- data.frame(prots_df_ordered[1],prots_df_ordered[2],prots_df_ordered[3],prots_df_ordered[4],
                         prots_df_ordered[9],prots_df_ordered[10],prots_df_ordered[11],prots_df_ordered[12])

prots_STEM <- data.frame(prots_df_ordered[5],prots_df_ordered[6],prots_df_ordered[7],prots_df_ordered[8],
                         prots_df_ordered[13],prots_df_ordered[14],prots_df_ordered[15],prots_df_ordered[16])

# Filter to keep only those proteins with at least 70% of non-NA values
prots_DIFF <- prots_DIFF[which(rowMeans(!is.na(prots_DIFF)) > 0.7), ]
prots_STEM <- prots_STEM[which(rowMeans(!is.na(prots_STEM)) > 0.7), ]
prots_all <- prots_df_ordered[which(rowMeans(!is.na(prots_STEM)) > 0.7), ]

# Create protein IDs vectors again with the proteins filtered in each group
protein_ids_DIFF <- rownames(prots_DIFF)
protein_ids_STEM <- rownames(prots_STEM)
protein_ids_all <- rownames(prots_all)

# Quick function to impute NAs with LoD/5

lod5_impute <- function(mat) {
  
  imputed_mat <- t(apply(mat, 1, function(row) {
   
     # Lowest detected value for this protein
    lod <- min(row, na.rm = TRUE)
    
    # Replace NAs with LoD/5
    row[is.na(row)] <- lod / 5
    
    return(row)
  }))
  
  # Preserve row/column names
  rownames(imputed_mat) <- rownames(mat)
  colnames(imputed_mat) <- colnames(mat)
  
  return(imputed_mat)
}

prots_DIFF <- as.data.frame(lod5_impute(prots_DIFF))
prots_STEM <- as.data.frame(lod5_impute(prots_STEM))
prots_all <- as.data.frame(lod5_impute(prots_all))

boxplot(prots_DIFF,outline=F,las=2,ylab="Protein Quantification (Differentiated)")
boxplot(prots_STEM,outline=F,las=2,ylab="Protein Quantification (Stem-like")
```

Variations between replicates of the same conditions show the presence of noise. We will use NormalyzerDE to choose the best normalization technique.

```{r}
## A data frame needs to be generated to specify the experimental design.
design_DIFF <- data.frame(sample=colnames(prots_DIFF),
                     group=c(rep("shNT_DIFF",4),rep("shRAD52_DIFF",4)))

design_STEM <- data.frame(sample=colnames(prots_STEM),
                     group=c(rep("shNT_STEM",4),rep("shRAD52_STEM",4)))

design_all <- data.frame(sample=colnames(prots_df_ordered),
                     group=c(rep("shNT_DIFF",4),
                             rep("shNT_STEM",4),
                             rep("shRAD52_DIFF",4),
                             rep("shRAD52_STEM",4)))

## The experimental design needs to be saved to a file.
write.table(x = design_DIFF, file = "normalyzer_design_DIFF.tsv",quote = F,
            row.names = F, sep = "\t")
write.table(x = design_STEM, file = "normalyzer_design_STEM.tsv",quote = F,
            row.names = F, sep = "\t")
write.table(x = design_all, file = "normalyzer_design_all.tsv",quote = F,
            row.names = F, sep = "\t")

# A table with the raw data is also saved. The values are increased +1 to make
# sure that all values are above 1, so that we can apply log transformations.
prots_DIFF_1 <- prots_DIFF + 1
write.table(x = prots_DIFF_1, file = "prots_DIFF.tsv", quote = F,
            row.names = F, sep = "\t")

prots_STEM_1 <- prots_STEM + 1
write.table(x = prots_STEM_1, file = "prots_STEM.tsv", quote = F,
            row.names = F, sep = "\t")

prots_all_1 <- prots_all + 1
write.table(x = prots_all_1, file = "prots_all.tsv", quote = F,
            row.names = F, sep = "\t")

## The following function runs all normalization techniques and generates
## evaluations measures for them. 
normalyzer(jobName = "prots_diff",designPath = "normalyzer_design_DIFF.tsv",
           dataPath = "prots_DIFF.tsv",outputDir = ".")

normalyzer(jobName = "prots_stem",designPath = "normalyzer_design_STEM.tsv",
           dataPath = "prots_STEM.tsv",outputDir = ".")

normalyzer(jobName = "prots_all",designPath = "normalyzer_design_all.tsv",
           dataPath = "prots_all.tsv",outputDir = ".")
```

We will pick Variance Stabilizing Normalization (VSN), as it seems to be reducing intragroup variance the best.

```{r}
normalized_prot_DIFF <- read.table(file="prots_diff/VSN-normalized.txt", header=T)
rownames(normalized_prot_DIFF) <- protein_ids_DIFF
boxplot(normalized_prot_DIFF,outline=F,las=2,ylab="Normalized Protein Quantification (Differentiated)")

normalized_prot_STEM <- read.table(file="prots_stem/VSN-normalized.txt", header=T)
rownames(normalized_prot_STEM) <- protein_ids_STEM
boxplot(normalized_prot_STEM,outline=F,las=2,ylab="Normalized Protein Quantification (Stem-like)")

normalized_protein <- read.table(file="prots_all/VSN-normalized.txt", header=T)
rownames(normalized_protein) <- protein_ids_all
boxplot(normalized_protein,outline=F,las=2,ylab="Normalized Protein Quantification (Stem-like)")
```

We can represent two different replicates in a scatter plot to check that they are mostly similar:

```{r}
plot(x = normalized_prot_DIFF$shNT_DIFF_1,
     y = normalized_prot_DIFF$shNT_DIFF_2,
     pch=19,col="grey",xlab="shNT_DIFF_1",ylab="shNT_DIFF_2",cex=0.5)
text(x=4,y=12,
     labels = paste(c(
      "cor = ",
      round(100*cor(normalized_prot_DIFF$shNT_DIFF_1,
                    normalized_prot_DIFF$shNT_DIFF_2),
            digits = 2),
      "%"), collapse=""))
```

We will analyze the data with Principal Component Analysis (PCA) in order to verify that no mistakes were made in labeling the samples and that samples with similar biological conditions cluster together as expected.

```{r}
# TransSTEMing the df for the function to work
pca_prot_all <- data.frame(colnames(normalized_protein),
                                  t(normalized_protein))
colnames(pca_prot_all)[1] <- "Sample"
pca_prot_all[1:8][1:6]

res_pca_all <- PCA(pca_prot_all, 
               graph = FALSE,
               scale.unit = TRUE,
               quali.sup = 1 )

fviz_pca_ind(res_pca_all, col.ind = design_all$group, 
             pointsize=2, pointshape=21,fill="black",
             repel = TRUE, 
             addEllipses = TRUE,ellipse.type = "confidence",
             legend.title="Conditions",
             title="",
             show_legend=TRUE,show_guide=TRUE)
```

The differences between cells with the Non-target short hairpin interference RNA (shNT) and the ones with target interference RNA against RAD52 (shRAD52) are clearly seen in DIFF (Glioblastoma Differentiated Cells), but not so much for the STEM (Glioblastoma Stem Cells).

We can also perform hierarchical clustering analysis over the PCA results and represent it in a dendrogram to check that the different conditions group up together as expected.

```{r}
res_hcpc_all <- HCPC(res_pca_all, graph=FALSE, nb.clust = 4)   

fviz_dend(res_hcpc_all,k=4,
          cex = 0.75,                       # Label size
          palette = "jco",               # Color palette see ?ggpubr::ggpar
          rect = TRUE, rect_fill = TRUE, # Add rectangle around groups
          rect_border = "jco",           # Rectangle color
          type="rectangle",
          labels_track_height = 700      # Augment the room for labels
)
```

As we saw in the PCA, the samples for STEM cells are very similar and the hierarchical clustering cannot differentiate between those with and without shRAD52. In the DIFF cells, the clustering differentiates better between the two conditions, but still mixes up some of the samples.

## DEGs

Differential expression analysis with limma:

### Individual limma contrasts

Contrasts between shRAD52 and shNT conditions will be made separately for STEM and DIFF cells:

```{r}
experimental_design_diff <- model.matrix(~ -1+factor(c(1,1,1,1,2,2,2,2)))
colnames(experimental_design_diff) <- c("shNT_DIFF","shRAD52_DIFF")
experimental_design_stem <- model.matrix(~ -1+factor(c(1,1,1,1,2,2,2,2)))
colnames(experimental_design_stem) <- c("shNT_STEM","shRAD52_STEM")

# normalized_prot_DIFF <- normalized_prot_DIFF[intersect_proteins,]

## Linear model fit to the experimental design, mean expression calculation
linear_fit_diff <- lmFit(normalized_prot_DIFF, experimental_design_diff)
linear_fit_stem <- lmFit(normalized_prot_STEM, experimental_design_stem)

## Contrast specification treatment-control
contrast_matrix_diff <- makeContrasts(shRAD52_DIFF-shNT_DIFF,
                                 levels=c("shNT_DIFF","shRAD52_DIFF"))
contrast_matrix_stem <- makeContrasts(shRAD52_STEM-shNT_STEM,
                                 levels=c("shNT_STEM","shRAD52_STEM"))

contrast_linear_fit_diff <- contrasts.fit(linear_fit_diff, contrast_matrix_diff)
contrast_results_diff <- eBayes(contrast_linear_fit_diff)
contrast_linear_fit_stem <- contrasts.fit(linear_fit_stem, contrast_matrix_stem)
contrast_results_stem <- eBayes(contrast_linear_fit_stem)
```

```{r}
## Results extraction for the first contrast (shRAD52_DIFF.shNT_DIFF) sorted according to logFC
shRAD52_DIFF.shNT_DIFF <- topTable(contrast_results_diff, 
                      number=nrow(normalized_prot_DIFF),
                      coef=1,
                      sort.by="logFC")
head(shRAD52_DIFF.shNT_DIFF)

## DEGs extraction
limma.log2fc <- shRAD52_DIFF.shNT_DIFF$logFC
limma.adjusted.pval <- shRAD52_DIFF.shNT_DIFF$adj.P.Val 
limma.pval <- shRAD52_DIFF.shNT_DIFF$P.Value 
limma.protein_ids <- rownames(shRAD52_DIFF.shNT_DIFF)

names(limma.log2fc) <- limma.protein_ids
names(limma.adjusted.pval) <- limma.protein_ids
names(limma.pval) <- limma.protein_ids

shRAD52_DIFF.shNT_DIFF.activated.genes <- limma.protein_ids[limma.log2fc > 0 & 
                                          limma.pval < 0.01]
shRAD52_DIFF.shNT_DIFF.repressed.genes <- limma.protein_ids[limma.log2fc < 0 & 
                                          limma.pval < 0.01]
length(shRAD52_DIFF.shNT_DIFF.activated.genes)
length(shRAD52_DIFF.shNT_DIFF.repressed.genes)

## Volcano plot
volcano_df <- data.frame(
  logFC = limma.log2fc,
  pval = limma.pval
)

volcano_df$negLogP <- -log10(volcano_df$pval)
volcano_df$gene <- rownames(volcano_df)

volcano_df$group <- "NS"
volcano_df$group[volcano_df$gene %in% shRAD52_DIFF.shNT_DIFF.activated.genes] <- "Up"
volcano_df$group[volcano_df$gene %in% shRAD52_DIFF.shNT_DIFF.repressed.genes] <- "Down"

# pick top genes for labeling (by significance)
top_genes <- volcano_df[order(volcano_df$pval), ][1:10, ]
max_fc <- max(abs(volcano_df$logFC), na.rm = TRUE) * 1.05

volcano <- ggplot(volcano_df, aes(x = logFC, y = negLogP)) +

  # background points (neutral grey, transparent)
  geom_point(
    color = "grey75",
    alpha = 0.6,
    size = 1.3
  ) +

  # highlighted groups
  geom_point(
    data = subset(volcano_df, group == "Up"),
    color = "#2C7BB6",
    alpha = 0.85,
    size = 1.5
  ) +
  geom_point(
    data = subset(volcano_df, group == "Down"),
    color = "#D7191C",
    alpha = 0.85,
    size = 1.5
  ) +

  # threshold lines (subtle, not distracting)
  geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = "grey60") +

  # label only top genes
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey50"
  ) +

  # axis + theme styling (this is what makes it “Nature-like”)
  labs(
    x = "log2 fold change",
    y = expression(-log[10]("p-value"))
  ) +
  
  coord_cartesian(xlim = c(-max_fc, max_fc),
                  ylim = c(0,6.5)) +

  theme_classic(base_size = 14) +
  theme(

    # remove minor clutter
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),

    # cleaner axis text
    axis.text = element_text(color = "black"),

    # no grid anywhere (very important for journal look)
    panel.grid = element_blank(),

    # slightly stronger axis titles
    axis.title = element_text(face = "bold"),

    # tight margins (optional but very “paper-like”)
    plot.margin = margin(10, 10, 10, 10)
  )

ggsave("volcano_RAD52DIFF.svg", plot = volcano, width = 6, height = 6)

################################ STEM RAD52 vs NT #########################

## Results extraction for the second contrast (shRAD52_STEM.shNT_STEM) sorted according to logFC
shRAD52_STEM.shNT_STEM <- topTable(contrast_results_stem, 
                      number=nrow(normalized_prot_STEM),
                      coef=1,
                      sort.by="logFC")
head(shRAD52_STEM.shNT_STEM)

## DEGs extraction
limma.log2fc <- shRAD52_STEM.shNT_STEM$logFC
limma.adjusted.pval <- shRAD52_STEM.shNT_STEM$adj.P.Val 
limma.pval <- shRAD52_STEM.shNT_STEM$P.Value 
limma.protein_ids <- rownames(shRAD52_STEM.shNT_STEM)

names(limma.log2fc) <- limma.protein_ids
names(limma.adjusted.pval) <- limma.protein_ids
names(limma.pval) <- limma.protein_ids

shRAD52_STEM.shNT_STEM.activated.genes <- limma.protein_ids[limma.log2fc > 0 & 
                                          limma.pval < 0.01]
shRAD52_STEM.shNT_STEM.repressed.genes <- limma.protein_ids[limma.log2fc < 0 & 
                                          limma.pval < 0.01]
length(shRAD52_STEM.shNT_STEM.activated.genes)
length(shRAD52_STEM.shNT_STEM.repressed.genes)

## Volcano plot
volcano_df <- data.frame(
  logFC = limma.log2fc,
  pval = limma.pval
)

volcano_df$negLogP <- -log10(volcano_df$pval)
volcano_df$gene <- rownames(volcano_df)

volcano_df$group <- "NS"
volcano_df$group[volcano_df$gene %in% shRAD52_STEM.shNT_STEM.activated.genes] <- "Up"
volcano_df$group[volcano_df$gene %in% shRAD52_STEM.shNT_STEM.repressed.genes] <- "Down"

# pick top genes for labeling (by significance)
top_genes <- volcano_df[order(volcano_df$pval), ][1:10, ]
max_fc <- max(abs(volcano_df$logFC), na.rm = TRUE) * 1.05

volcano <- ggplot(volcano_df, aes(x = logFC, y = negLogP)) +

  # background points (neutral grey, transparent)
  geom_point(
    color = "grey75",
    alpha = 0.6,
    size = 1.3
  ) +

  # highlighted groups
  geom_point(
    data = subset(volcano_df, group == "Up"),
    color = "#2C7BB6",
    alpha = 0.85,
    size = 1.5
  ) +
  geom_point(
    data = subset(volcano_df, group == "Down"),
    color = "#D7191C",
    alpha = 0.85,
    size = 1.5
  ) +

  # threshold lines (subtle, not distracting)
  geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = "grey60") +

  # label only top genes
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey50"
  ) +

  # axis + theme styling (this is what makes it “Nature-like”)
  labs(
    x = "log2 fold change",
    y = expression(-log[10]("p-value"))
  ) +
  
  coord_cartesian(xlim = c(-max_fc, max_fc)) +

  theme_classic(base_size = 14) +
  theme(

    # remove minor clutter
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),

    # cleaner axis text
    axis.text = element_text(color = "black"),

    # no grid anywhere (very important for journal look)
    panel.grid = element_blank(),

    # slightly stronger axis titles
    axis.title = element_text(face = "bold"),

    # tight margins (optional but very “paper-like”)
    plot.margin = margin(10, 10, 10, 10)
  )

ggsave("volcano_RAD52STEM.svg", plot = volcano, width = 6, height = 6)
```

### Grouped limma analysis

For the comparison between STEM and DIFF cells, there is no need to separate both types of cells in different limma models.

```{r}
## Experimental design specification
experimental.design <- model.matrix(~ -1+factor(c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,4)))
colnames(experimental.design) <- c("shNT_DIFF","shNT_STEM","shRAD52_DIFF","shRAD52_STEM")

## Linear model fit to the experimental design, mean expression calculation
linear.fit <- lmFit(normalized_protein, experimental.design)

## Contrast specification treatment-control
contrast.matrix <- makeContrasts(shRAD52_DIFF-shNT_DIFF,shRAD52_STEM-shNT_STEM,shRAD52_STEM-shRAD52_DIFF,shNT_STEM-shNT_DIFF,
                                 levels=c("shNT_DIFF","shNT_STEM","shRAD52_DIFF","shRAD52_STEM"))

contrast.linear.fit <- contrasts.fit(linear.fit, contrast.matrix)
contrast.results <- eBayes(contrast.linear.fit, robust = TRUE)
# Trying with robust=TRUE to try and make it less sensible to variance imbalance
```

```{r}
## Results extraction for the fourth contrast (coef=4) sorted according to logFC
shNT_STEM.shNT_DIFF <- topTable(contrast.results, 
                      number=nrow(normalized_protein),
                      coef=4,
                      sort.by="logFC")
head(shNT_STEM.shNT_DIFF)

## DEGs extraction
limma.log2fc <- shNT_STEM.shNT_DIFF$logFC
limma.adjusted.pval <- shNT_STEM.shNT_DIFF$adj.P.Val
limma.pval <- shNT_STEM.shNT_DIFF$P.Value 
limma.protein_ids <- rownames(shNT_STEM.shNT_DIFF)

names(limma.log2fc) <- limma.protein_ids
names(limma.adjusted.pval) <- limma.protein_ids
names(limma.pval) <- limma.protein_ids

shNT_STEM.shNT_DIFF.activated.genes <- limma.protein_ids[limma.log2fc > 0.58 & 
                                          limma.adjusted.pval < 0.01]
shNT_STEM.shNT_DIFF.repressed.genes <- limma.protein_ids[limma.log2fc < - 0.58 & 
                                          limma.adjusted.pval < 0.01]
length(shNT_STEM.shNT_DIFF.activated.genes)
length(shNT_STEM.shNT_DIFF.repressed.genes)

## Volcano plot
volcano_df <- data.frame(
  logFC = limma.log2fc,
  padj = limma.adjusted.pval
)

volcano_df$negLogP <- -log10(volcano_df$padj)
volcano_df$gene <- rownames(volcano_df)

volcano_df$group <- "NS"
volcano_df$group[volcano_df$gene %in% shNT_STEM.shNT_DIFF.activated.genes] <- "Up"
volcano_df$group[volcano_df$gene %in% shNT_STEM.shNT_DIFF.repressed.genes] <- "Down"

# pick top genes for labeling (by significance)
top_genes <- volcano_df[order(volcano_df$padj), ][1:10, ]
max_fc <- max(abs(volcano_df$logFC), na.rm = TRUE) * 1.05

volcano <- ggplot(volcano_df, aes(x = logFC, y = negLogP)) +

  # background points (neutral grey, transparent)
  geom_point(
    color = "grey75",
    alpha = 0.6,
    size = 1.3
  ) +

  # highlighted groups
  geom_point(
    data = subset(volcano_df, group == "Up"),
    color = "#2C7BB6",
    alpha = 0.85,
    size = 1.5
  ) +
  geom_point(
    data = subset(volcano_df, group == "Down"),
    color = "#D7191C",
    alpha = 0.85,
    size = 1.5
  ) +

  # threshold lines (subtle, not distracting)
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey60") +
  geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = "grey60") +

  # label only top genes
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey50"
  ) +

  # axis + theme styling (this is what makes it “Nature-like”)
  labs(
    x = "log2 fold change",
    y = expression(-log[10]("adjusted p-value"))
  ) +
  
  coord_cartesian(xlim = c(-max_fc, max_fc)) +

  theme_classic(base_size = 14) +
  theme(

    # remove minor clutter
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),

    # cleaner axis text
    axis.text = element_text(color = "black"),

    # no grid anywhere (very important for journal look)
    panel.grid = element_blank(),

    # slightly stronger axis titles
    axis.title = element_text(face = "bold"),

    # tight margins (optional but very “paper-like”)
    plot.margin = margin(10, 10, 10, 10)
  )

ggsave("volcano_STEMDIFF.svg", plot = volcano, width = 6, height = 6)
```


## Enrichment

### GO Enrichment

With the first comparison shRAD52_DIFF vs shNT_DIFF (activated and repressed genes). Even with a pvalueCutoff of 0.1, there are almost no enriched terms (there are 2 of them in the repressed genes, related to mithocondrial membrane)

```{r}
activated.enrich.go.shRAD52_DIFF.shNT_DIFF <- enrichGO(gene = shRAD52_DIFF.shNT_DIFF.activated.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.1,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(activated.enrich.go.shRAD52_DIFF.shNT_DIFF)
#treeplot(pairwise_termsim(activated.enrich.go.shRAD52_DIFF.shNT_DIFF),showCategory = 20)

repressed.enrich.go.shRAD52_DIFF.shNT_DIFF <- enrichGO(gene = shRAD52_DIFF.shNT_DIFF.repressed.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.1,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(repressed.enrich.go.shRAD52_DIFF.shNT_DIFF)
#treeplot(pairwise_termsim(repressed.enrich.go.shRAD52_DIFF.shNT_DIFF),showCategory = 20)

```

Now, for shRAD52_STEM vs shNT_STEM

```{r}
activated.enrich.go.shRAD52_STEM.shNT_STEM <- enrichGO(gene = shRAD52_STEM.shNT_STEM.activated.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.1,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(activated.enrich.go.shRAD52_STEM.shNT_STEM)
#treeplot(pairwise_termsim(activated.enrich.go.shRAD52_DIFF.shNT_DIFF),showCategory = 20)

repressed.enrich.go.shRAD52_STEM.shNT_STEM <- enrichGO(gene = shRAD52_STEM.shNT_STEM.repressed.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.1,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(repressed.enrich.go.shRAD52_STEM.shNT_STEM)
#treeplot(pairwise_termsim(repressed.enrich.go.shRAD52_DIFF.shNT_DIFF),showCategory = 20)
```

And again with shNT_STEM vs shNT_DIFF:

```{r}
activated.enrich.go.shNT_STEM.shNT_DIFF <- enrichGO(gene = shNT_STEM.shNT_DIFF.activated.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.5,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(activated.enrich.go.shNT_STEM.shNT_DIFF)
#treeplot(pairwise_termsim(activated.enrich.go.shNT_STEM.shNT_DIFF),showCategory = 20)
#barplot(activated.enrich.go.shNT_STEM.shNT_DIFF, showCategory = 20)

repressed.enrich.go.shNT_STEM.shNT_DIFF <- enrichGO(gene = shNT_STEM.shNT_DIFF.repressed.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.5,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(repressed.enrich.go.shNT_STEM.shNT_DIFF)
# treeplot(pairwise_termsim(repressed.enrich.go.shNT_STEM.shNT_DIFF),showCategory = 20)
# barplot(repressed.enrich.go.shNT_STEM.shNT_DIFF, showCategory = 20)
```

### KEGGs

Now, we will try to obtain enriched KEGG pathways for the shRAD52_DIFF-shNT_DIFF comparison:

```{r}
activated.df <- bitr(
    shRAD52_DIFF.shNT_DIFF.activated.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

repressed.df <- bitr(
    shRAD52_DIFF.shNT_DIFF.repressed.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

universe.df <- bitr(
    protein_ids,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

activated.enrich.kegg.shRAD52_DIFF.shNT_DIFF <- enrichKEGG(
    gene = activated.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- activated.enrich.kegg.shRAD52_DIFF.shNT_DIFF@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

repressed.enrich.kegg.shRAD52_DIFF.shNT_DIFF <- enrichKEGG(
    gene = repressed.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- repressed.enrich.kegg.shRAD52_DIFF.shNT_DIFF@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)
```

For the shRAD52_STEM-shNT_STEM comparison:

```{r}
activated.df <- bitr(
    shRAD52_STEM.shNT_STEM.activated.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

repressed.df <- bitr(
    shRAD52_STEM.shNT_STEM.repressed.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

universe.df <- bitr(
    protein_ids,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

activated.enrich.kegg.shRAD52_STEM.shNT_STEM <- enrichKEGG(
    gene = activated.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- activated.enrich.kegg.shRAD52_STEM.shNT_STEM@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

repressed.enrich.kegg.shRAD52_STEM.shNT_STEM <- enrichKEGG(
    gene = repressed.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- repressed.enrich.kegg.shRAD52_STEM.shNT_STEM@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

```

For the record, the shNT_STEM-shNT_DIFF comparison:

```{r}
activated.df <- bitr(
    shNT_STEM.shNT_DIFF.activated.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

repressed.df <- bitr(
    shNT_STEM.shNT_DIFF.repressed.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

universe.df <- bitr(
    protein_ids,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

activated.enrich.kegg.shNT_STEM.shNT_DIFF <- enrichKEGG(
    gene = activated.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- activated.enrich.kegg.shNT_STEM.shNT_DIFF@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

repressed.enrich.kegg.shNT_STEM.shNT_DIFF <- enrichKEGG(
    gene = repressed.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- repressed.enrich.kegg.shNT_STEM.shNT_DIFF@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)
```

## GSEA (Gene Set Enrichment Analysis)

Gene Set Enrichment Analysis (GSEA) is a computational method that determines whether an a priori defined set of genes shows statistically significant, concordant differences between two biological states

### fgsea()

Another way of obtaining enriched pathways. We will first try with fgsea(). First, the differentiated cells shRAD52_DIFF vs shNT_DIFF:

```{r}
set.seed(123) # Setting a fixed seed, because GSEA calculates the pvalues with
              # Monte-Carlo sampling and so they may vary

# Loading Hallmark sets from the Molecular Signatures Database (MSigDB)
msig <- msigdbr(species = "Homo sapiens", collection = "H")  

# Selecting only the gene sets with their corresponding genes names
gene_sets <- msig %>%
  dplyr::select(gs_name, gene_symbol) %>%
  split(x = .$gene_symbol, f = .$gs_name)

# Using the t-statistic to rank the genes
ranks <- shRAD52_DIFF.shNT_DIFF$t
names(ranks) <- rownames(shRAD52_DIFF.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

fgseaRes.shRAD52_DIFF.shNT_DIFF <- fgsea(
  pathways = gene_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)
```

There seems to be some pathways weakly enriched (only pvalue is <0.05 for some, but no adjusted pvalue). We will try the other comparisons:

```{r}
## RAD52_STEM vs NT_STEM
ranks <- shRAD52_STEM.shNT_STEM$t
names(ranks) <- rownames(shRAD52_STEM.shNT_STEM)

ranks <- sort(ranks, decreasing = TRUE)

fgseaRes.shRAD52_STEM.shNT_STEM <- fgsea(
  pathways = gene_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

###################

## STEM vs DIFF
ranks <- shNT_STEM.shNT_DIFF$t
names(ranks) <- rownames(shNT_STEM.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

fgseaRes.shNT_STEM.shNT_DIFF <- fgsea(
  pathways = gene_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)
```

#### FGSEA Plots

For shNT_STEM.shNT_DIFF:

```{r}
## Enrichment plot
ranks <- shNT_STEM.shNT_DIFF$t
names(ranks) <- rownames(shNT_STEM.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

plotEnrichment(gene_sets[["HALLMARK_DNA_REPAIR"]], ranks)

## Dot plot

dotplot <- ggplot(plot_df,
       aes(NES, reorder(pathway, NES))) +
    geom_point(aes(size = size,
                   color = padj))

ggsave("Dotplot_STEMDIFF.svg", plot = dotplot, width = 8, height = 6)
```

For shRAD52_STEM.shNT_STEM and shRAD52_DIFF.shNT_DIFF:

```{r}
# We will represent those with a pval < 0.05, since the adjusted pval
# is too restrictive for this comparison
plot_df <- fgseaRes.shRAD52_STEM.shNT_STEM %>%
  filter(pval < 0.05) %>%
  arrange(NES)

dotplot <- ggplot(plot_df,
       aes(NES, reorder(pathway, NES))) +
    geom_point(aes(size = size,
                   color = pval))

ggsave("Dotplot_RAD52STEM.svg", plot = dotplot, width = 8, height = 6)

################

plot_df <- fgseaRes.shRAD52_DIFF.shNT_DIFF %>%
  filter(pval < 0.05) %>%
  arrange(NES)

dotplot <- ggplot(plot_df,
       aes(NES, reorder(pathway, NES))) +
    geom_point(aes(size = size,
                   color = pval))

ggsave("Dotplot_RAD52DIFF.svg", plot = dotplot, width = 8, height = 6)
```

### camera()

Now I will try with camera(), part of the limma package, to check complementary enrichment methods:

```{r}
cameraRes.shRAD52_DIFF.shNT_DIFF <- camera(
    y = normalized_prot_DIFF,
    index = gene_sets,
    design = experimental_design_diff,
    contrast = contrast_matrix_diff[,1])

cameraRes.shRAD52_STEM.shNT_STEM <- camera(
    y = normalized_prot_STEM,
    index = gene_sets,
    design = experimental_design_stem,
    contrast = contrast_matrix_stem[,1])

cameraRes.shNT_STEM.shNT_DIFF <- camera(
    y = normalized_protein,
    index = gene_sets,
    design = experimental.design,
    contrast = contrast.matrix[,4])
```

### fry()

Yet another method for enrichment of gene sets, fry():

```{r}
fryRes.shRAD52_DIFF.shNT_DIFF <- fry(
  normalized_prot_DIFF,
  index = gene_sets,
  design = experimental_design_diff,
  contrast = contrast_matrix_diff[,1])

fryRes.shRAD52_STEM.shNT_STEM <- fry(
  normalized_prot_STEM,
  index = gene_sets,
  design = experimental_design_stem,
  contrast = contrast_matrix_stem[,1])

fryRes.shNT_STEM.shNT_DIFF <- fry(
  normalized_protein,
  index = gene_sets,
  design = experimental.design,
  contrast = contrast.matrix[,4])
```

## GSVA (Gene Set Variation Analysis)

Gene Set Variation Analysis (GSVA) is a GSE method that estimates variation of pathway activity over a sample population in an unsupervised manner.

```{r}
expr_mat <- as.matrix(normalized_protein)

param <- gsvaParam(
  exprData = expr_mat,
  geneSets = gene_sets,
  kcdf = "Gaussian"
)

gsva_res <- gsva(param)

## Linear model fit to the experimental design, mean expression calculation
gsva.linear.fit <- lmFit(gsva_res, experimental.design)

## Contrast specification treatment-control
gsva.contrast.matrix <- makeContrasts(shRAD52_DIFF-shNT_DIFF,shRAD52_STEM-shNT_STEM,
                                      shRAD52_STEM-shRAD52_DIFF,shNT_STEM-shNT_DIFF,
                                 levels=c("shNT_DIFF","shNT_STEM","shRAD52_DIFF","shRAD52_STEM"))

gsva.contrast.linear.fit <- contrasts.fit(gsva.linear.fit, gsva.contrast.matrix)
gsva.contrast.results <- eBayes(gsva.contrast.linear.fit)

## Results extraction for the first contrast shRAD52_DIFF - shNT_DIFF) sorted according to logFC
grouped_gsva.shRAD52_DIFF.shNT_DIFF <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res),
                      coef=1,
                      sort.by="P")
head(grouped_gsva.shRAD52_DIFF.shNT_DIFF)

## Results extraction for the second contrast (shRAD52_STEM - shNT_STEM) sorted according to logFC
grouped_gsva.shRAD52_STEM.shNT_STEM <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res),
                      coef=2,
                      sort.by="logFC")
head(grouped_gsva.shRAD52_STEM.shNT_STEM)

## Results extraction for the fourth contrast (shNT_STEM - shNT_DIFF) sorted according to logFC
grouped_gsva.shNT_STEM.shNT_DIFF <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res),
                      coef=4,
                      sort.by="logFC")
head(grouped_gsva.shNT_STEM.shNT_DIFF)
```

This meta structure here created is used for the boxplot representations.

```{r}
## Experimental design specification
samples <- colnames(normalized_protein)

# Create metadata
meta <- data.frame(
  sample_id = samples,
  shRNA = ifelse(grepl("^shNT", samples), "Control", "RAD52"),
  cell_type = ifelse(grepl("_DIFF_", samples), "DGC", "GSC"),
  row.names = samples
)

# Convert to factors with correct reference levels
meta$cell_type <- factor(meta$cell_type, levels = c("DGC", "GSC"))
meta$shRNA <- factor(meta$shRNA, levels = c("Control", "RAD52"))

# Making sure the metadata rows have the same order as the columns in the GSVA
meta <- meta[colnames(gsva_res), ]
```

### GSVA Boxplots

Plotting boxplots of relevant pathway, representing the GSVA score of each condition:

```{r}
df_dna_repair <- data.frame(
  score = gsva_res["HALLMARK_DNA_REPAIR", ],
  meta
)
ggplot(df_dna_repair, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_DNA_REPAIR")

df_e2f_targets <- data.frame(
  score = gsva_res["HALLMARK_E2F_TARGETS", ],
  meta
)
ggplot(df_e2f_targets, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_E2F_TARGETS")

df_jak_stat3 <- data.frame(
  score = gsva_res["HALLMARK_IL6_JAK_STAT3_SIGNALING", ],
  meta
)
ggplot(df_jak_stat3, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_IL6_JAK_STAT3_SIGNALING")

df_myc_targets <- data.frame(
  score = gsva_res["HALLMARK_MYC_TARGETS_V2", ],
  meta
)
ggplot(df_myc_targets, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_MYC_TARGETS_V2")

df_epithelial_mesenchymal <- data.frame(
  score = gsva_res["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION", ],
  meta
)
gsva_plot <- ggplot(df_epithelial_mesenchymal,
                    aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot(width = 0.5, position = position_dodge(width = 0.7)) +
  theme_minimal() +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  ggtitle("HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION")

ggsave("GSVA_EMT.svg", plot = gsva_plot, width = 6, height = 6)

df_mtor <- data.frame(
  score = gsva_res["HALLMARK_PI3K_AKT_MTOR_SIGNALING", ],
  meta
)
ggplot(df_mtor, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_PI3K_AKT_MTOR_SIGNALING")

df_oxphos <- data.frame(
  score = gsva_res["HALLMARK_OXIDATIVE_PHOSPHORYLATION", ],
  meta
)
ggplot(df_oxphos, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_OXIDATIVE_PHOSPHORYLATION")

df_uv_up <- data.frame(
  score = gsva_res["HALLMARK_UV_RESPONSE_UP", ],
  meta
)
gsva_plot <- ggplot(df_uv_up,
                    aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot(width = 0.5, position = position_dodge(width = 0.7)) +
  theme_minimal() +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  ggtitle("HALLMARK_UV_RESPONSE_UP")

ggsave("GSVA_UV_UP.svg", plot = gsva_plot, width = 6, height = 6)

df_apoptosis <- data.frame(
  score = gsva_res["HALLMARK_APOPTOSIS", ],
  meta
)
gsva_plot <- ggplot(df_apoptosis,
                    aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot(width = 0.5, position = position_dodge(width = 0.7)) +
  theme_minimal() +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  ggtitle("GSVA_APOPTOSIS.svg")

ggsave("GSVA_APOPTOSIS.svg", plot = gsva_plot, width = 6, height = 6)

df_hypoxia <- data.frame(
  score = gsva_res["HALLMARK_HYPOXIA", ],
  meta
)
ggplot(df_hypoxia, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_HYPOXIA")

df_estrogen_late <- data.frame(
  score = gsva_res["HALLMARK_ESTROGEN_RESPONSE_LATE", ],
  meta
)
ggplot(df_estrogen_late, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_ESTROGEN_RESPONSE_LATE")
```

## Stemness

We are going to try to measure the "stemness" of the differents groups. For this, we will use different gene sets that are related to glioblastoma stem cells. These sets come from 3 different articles:

-   A CD133-related gene expression signature identifies an aggressive glioblastoma subtype with excessive mutations. <https://pmc.ncbi.nlm.nih.gov/articles/PMC3029739/>
- An embryonic stem cell-like gene expression signature in poorly differentiated aggressive human tumors. <https://pubmed.ncbi.nlm.nih.gov/18443585/>
- Module map of stem cell genes guides creation of epithelial cancer stem cells. <https://pubmed.ncbi.nlm.nih.gov/18397753/>

Loading the data from the paper:

```{r}
# 1. CD133+ glioblastoma stem cell signature (core subset)
GBM_CD133_STEMNESS = c(
  "AKAP4", "ARHGAP11A", "ASPM", "BARD1", "BIRC5", "BRCA1", "C12orf32", 
  "C17orf80", "C2orf48", "C4orf21", "CAPN14", "CASC5", "CCDC102A", 
  "CCDC111", "CCDC15", "CDCA2", "CDKN3", "CENPH", "CENPK", "CKAP2L", 
  "CKS2", "CTNNAL1", "DHFR", "DHX57", "DIAPH3", "DLGAP5", "DTL", 
  "DTYMK", "ECT2", "ENAH", "FANCI", "FBXO5", "GGH", "GINS2", "GMNN", 
  "H2AFZ", "HMGB2", "IFNA17", "IFNA4", "JAM2", "KIAA0101", "KIF11", 
  "KIF15", "KIF2C", "KIF4A", "KNTC1", "LIG1", "LMAN1L", "LOC91431", 
  "MAD2L1", "MCM2", "MCM3", "MELK", "MND1", "MORN2", "NACA", "NCAPH", 
  "NDC80", "NEK2", "NMU", "NUF2", "PBK", "PCNA", "POLQ", "PRIM1", 
  "PROM1", "PSG5", "PTTG1", "PTTG3", "PXMP2", "RAD51", "RANBP1", 
  "RRM2", "RTKN", "SGOL1", "SLC2A11", "SMC2", "SNRPE", "SYTL4", 
  "TIMELESS", "TM4SF1", "TMEM106C", "TOP2A", "TPX2", "TRIP13", 
  "TROAP", "TTK", "TYMS", "WDR34")

# Filter to use only the genes in our dataset
GBM_CD133_STEMNESS <- 
  GBM_CD133_STEMNESS[GBM_CD133_STEMNESS %in% rownames(normalized_protein)]

wong_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "WONG_EMBRYONIC_STEM_CELL_CORE") %>%
  pull(gene_symbol)

benporath_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "BENPORATH_ES_1") %>%
  pull(gene_symbol)

gbm_stemness_sets <- list(GBM_CD133_STEMNESS,
                          wong_stem_markers, benporath_stem_markers
                          )

names(gbm_stemness_sets) <- c(
    "Stemness_CD133_signature",
    "Wong Stem Markers",
    "Benporath Stem Markers"
    )
```

We can check how many of the genes in the gene sets are actually in our proteomic data:

```{r}
lapply(gbm_stemness_sets, function(gs) {
  sum(gs %in% rownames(normalized_protein))
})
```

Calculating stemness with GSEA

```{r}
# RAD52 DIFF
ranks <- shRAD52_DIFF.shNT_DIFF$t
names(ranks) <- rownames(shRAD52_DIFF.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

stemness.fgseaRes.shRAD52_DIFF.shNT_DIFF <- fgsea(
  pathways = gbm_stemness_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

stemness.fgseaRes.shRAD52_DIFF.shNT_DIFF

# RAD52 STEM
ranks <- shRAD52_STEM.shNT_STEM$t
names(ranks) <- rownames(shRAD52_STEM.shNT_STEM)

ranks <- sort(ranks, decreasing = TRUE)

stemness.fgseaRes.shRAD52_STEM.shNT_STEM <- fgsea(
  pathways = gbm_stemness_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

stemness.fgseaRes.shRAD52_STEM.shNT_STEM

# STEM - DIFF
ranks <- shNT_STEM.shNT_DIFF$t
names(ranks) <- rownames(shNT_STEM.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

stemness.fgseaRes.shNT_STEM.shNT_DIFF <- fgsea(
  pathways = gbm_stemness_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

stemness.fgseaRes.shNT_STEM.shNT_DIFF
```

## Transcription Factor Enrichment

### TF Enrichment Calculation

We will try infer the activated Transcription Factors (TF) from the protein expression dataset using Dorothea and decoupleR

```{r}
# Using the t-statistic as a measure of change
stat_vec <- shNT_STEM.shNT_DIFF$t
names(stat_vec) <- rownames(shNT_STEM.shNT_DIFF)

# Loading the TF data from dorothea
data(dorothea_hs, package = "dorothea")

# Keeping only higher-confidence interactions (dorothea classifies the
# confidence of the interactions in 5 categories, A-E):
regulons_AB <- dorothea_hs %>%
  filter(confidence %in% c("A","B"))

# Creating an input matrix for decoupleR
mat <- matrix(
  stat_vec,
  ncol = 1,
  dimnames = list(names(stat_vec), "STEM_vs_DIFF"))

# We will run several models and see the consensus features 
TF_ulm_STEM_DIFF <- run_ulm(
  mat = mat,
  network = regulons_AB,
  .source = "tf",
  .target = "target",
  .mor = "mor")

TF_mlm_STEM_DIFF <- run_mlm(
  mat = mat,
  network = regulons_AB,
  .source = "tf",
  .target = "target",
  .mor = "mor")

# Now, to select only those which are significant in both methods, for higher
# confidence
significant_TF_mlm_STEM_DIFF <- TF_mlm_STEM_DIFF[TF_mlm_STEM_DIFF$p_value<0.05,]
significant_TF_ulm_STEM_DIFF <- TF_ulm_STEM_DIFF[TF_ulm_STEM_DIFF$p_value<0.05,]

intersection_TF_STEM_DIFF <- intersect(significant_TF_mlm_STEM_DIFF$source,significant_TF_ulm_STEM_DIFF$source)
```

For the RAD52 vs NT comparison in STEM cells:

```{r}
# Using the t-statistic as a measure of change
stat_vec <- shRAD52_STEM.shNT_STEM$t
names(stat_vec) <- rownames(shRAD52_STEM.shNT_STEM)

# Loading the TF data from dorothea
data(dorothea_hs, package = "dorothea")

# Keeping only higher-confidence interactions (dorothea classifies the
# confidence of the interactions in 5 categories, A-E):
regulons_AB <- dorothea_hs %>%
  filter(confidence %in% c("A","B"))

# Creating an input matrix for decoupleR
mat <- matrix(
  stat_vec,
  ncol = 1,
  dimnames = list(names(stat_vec), "RAD52_vs_NT"))

# We will run several models and see the consensus features 
TF_ulm_RAD52_STEM <- run_ulm(
  mat = mat,
  network = regulons_AB,
  .source = "tf",
  .target = "target",
  .mor = "mor")

TF_mlm_RAD52_STEM <- run_mlm(
  mat = mat,
  network = regulons_AB,
  .source = "tf",
  .target = "target",
  .mor = "mor")

# Now, to select only those which are significant in both methods, for higher
# confidence
significant_TF_mlm_RAD52_STEM <- TF_mlm_RAD52_STEM[TF_mlm_RAD52_STEM$p_value<0.05,]
significant_TF_ulm_RAD52_STEM <- TF_ulm_RAD52_STEM[TF_ulm_RAD52_STEM$p_value<0.05,]

intersection_TF_RAD52_STEM <- intersect(significant_TF_mlm_RAD52_STEM$source,significant_TF_ulm_RAD52_STEM$source)
```

And for the RAD52 vs NT comparison in DIFF cells:

```{r}
# Using the t-statistic as a measure of change
stat_vec <- shRAD52_DIFF.shNT_DIFF$t
names(stat_vec) <- rownames(shRAD52_DIFF.shNT_DIFF)

# Loading the TF data from dorothea
data(dorothea_hs, package = "dorothea")

# Keeping only higher-confidence interactions (dorothea classifies the
# confidence of the interactions in 5 categories, A-E):
regulons_ABC <- dorothea_hs %>%
  filter(confidence %in% c("A","B","C"))  # Including category C because of lack
                                          # of results using only A and B

# Creating an input matrix for decoupleR
mat <- matrix(
  stat_vec,
  ncol = 1,
  dimnames = list(names(stat_vec), "RAD52_vs_NT"))

# We will run several models and see the consensus features 
TF_ulm_RAD52_DIFF <- run_ulm(
  mat = mat,
  network = regulons_ABC,
  .source = "tf",
  .target = "target",
  .mor = "mor")

TF_mlm_RAD52_DIFF <- run_mlm(
  mat = mat,
  network = regulons_ABC,
  .source = "tf",
  .target = "target",
  .mor = "mor")

# Now, to select only those which are significant in both methods, for higher
# confidence
significant_TF_mlm_RAD52_DIFF <- TF_mlm_RAD52_DIFF[TF_mlm_RAD52_DIFF$p_value<0.05,]
significant_TF_ulm_RAD52_DIFF <- TF_ulm_RAD52_DIFF[TF_ulm_RAD52_DIFF$p_value<0.05,]
intersection_TF_RAD52_DIFF <- intersect(significant_TF_mlm_RAD52_DIFF$source,significant_TF_ulm_RAD52_DIFF$source)
```

Dotplot and barplot visualization:

```{r}
## Visualization

target_count <- regulons_ABC %>%
  filter(target %in% rownames(mat)) %>%
  count(tf, name = "n_targets")

plot_df <- TF_mlm_RAD52_DIFF %>%
  left_join(target_count,
            by = c("source" = "tf")) %>%
  mutate(
    neglog10p = -log10(p_value),
    Significant = p_value < 0.05) %>%
  filter(p_value < 0.05)

plot_df$Label <- ifelse(
    plot_df$p_value < 0.02 &
    abs(plot_df$score) > 2,
    plot_df$source,
    ""
)

ggplot(plot_df,
       aes(score,
           reorder(source, score))) +

    geom_point(aes(size = n_targets,
                   colour = neglog10p),
               alpha = 0.9) +

    geom_text_repel(aes(label = Label),
                    size = 3.5,
                    max.overlaps = Inf) +

    scale_colour_gradient(
        low = "grey80",
        high = "#D73027",
        name = expression(-log[10](italic(P)))
    ) +

    scale_size(
        range = c(2,10),
        name = "Detected\ntargets"
    ) +

    geom_vline(xintercept = 0,
               linetype = 2,
               colour = "grey60") +

    labs(x = "TF activity score (ULM)",
         y = NULL) +

    theme_classic(base_size = 14) +

    theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "right"
    )

# Barplot
ggplot(plot_df,
       aes(x = score,
           y = reorder(source, score))) +
    geom_col() +
    theme_classic()
```

Consensus MLM+ULM plot (RAD52_DIFF):

```{r}
target_count <- regulons_ABC %>%
  filter(target %in% rownames(mat)) %>%
  count(tf, name = "n_targets")

consensus_df <- TF_ulm_RAD52_DIFF %>%
  dplyr::select(source,
         ulm_score = score,
         ulm_p = p_value) %>%
  inner_join(
    TF_mlm_RAD52_DIFF %>%
      dplyr::select(source,
             mlm_score = score,
             mlm_p = p_value),
    by = "source"
  ) %>%
  left_join(
    target_count,
    by = c("source" = "tf")
  )

consensus_df <- consensus_df %>%
  mutate(
    consensus_p = pmax(ulm_p, mlm_p),      # the less significant p-value
    neglog10p = -log10(consensus_p),

    significant =
      ulm_p < 0.05 &
      mlm_p < 0.05,

    label = ifelse(significant, source, "")
  )

consensus_df <- consensus_df %>%
  mutate(mean_score = (ulm_score + mlm_score)/2)

consensus_rad52diff_plot <- ggplot(consensus_df,
       aes(x = ulm_score,
           y = mlm_score,
           colour = mean_score)) +
  
    # --- quadrant shading ---
  annotate("rect",
           xmin = -Inf, xmax = 0,
           ymin = -Inf, ymax = 0,
           fill = "#3B6FB6", alpha = 0.08) +

  annotate("rect",
           xmin = 0, xmax = Inf,
           ymin = 0, ymax = Inf,
           fill = "#C0392B", alpha = 0.08) +

  ################
  
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "grey80") +

  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey80") +

  geom_abline(intercept = 0,
              slope = 1,
              colour = "grey60",
              linewidth = 0.5) +

  geom_point(
    aes(size = n_targets,
        colour = neglog10p),
    alpha = 0.85
  ) +

  geom_text_repel(
  aes(label = label),
  fontface = "bold",
  size = 4,
  box.padding = 0.5,
  point.padding = 0.4,
  segment.color = "grey60",
  segment.size = 0.4,
  max.overlaps = Inf
  ) +

  scale_colour_gradient2(
    low = "blue",
    mid = "grey92",
    high = "red",
    midpoint = 0
  ) +

  scale_size(
    range = c(2,10),
    name = "Detected\ntargets"
  ) +

  labs(
    x = "ULM activity score",
    y = "MLM activity score"
  ) +

  #coord_equal() +
  
  coord_cartesian(xlim = c(-3.2, 3.4), ylim = c(-2.9, 3.4)) +
  
  geom_point(
    data = subset(consensus_df, significant),
    shape = 21,
    stroke = 1.2,
    colour = "black",
    fill = NA,
    aes(size = n_targets)
  ) +

  theme_classic(base_size = 14)

ggsave("TF_RAD52DIFF.svg", plot = consensus_rad52diff_plot, width = 8, height = 6)
```

Consensus MLM+ULM plot (RAD52_STEM)

```{r}
target_count <- regulons_AB %>%
  filter(target %in% rownames(mat)) %>%
  count(tf, name = "n_targets")

consensus_df <- TF_ulm_RAD52_STEM %>%
  dplyr::select(source,
         ulm_score = score,
         ulm_p = p_value) %>%
  inner_join(
    TF_mlm_RAD52_STEM %>%
      dplyr::select(source,
             mlm_score = score,
             mlm_p = p_value),
    by = "source"
  ) %>%
  left_join(
    target_count,
    by = c("source" = "tf")
  )

consensus_df <- consensus_df %>%
  mutate(
    consensus_p = pmax(ulm_p, mlm_p),      # the less significant p-value
    neglog10p = -log10(consensus_p),

    significant =
      ulm_p < 0.05 &
      mlm_p < 0.05,

    label = ifelse(significant, source, "")
  )

consensus_df <- consensus_df %>%
  mutate(mean_score = (ulm_score + mlm_score)/2)

consensus_rad52stem_plot <- ggplot(consensus_df,
       aes(x = ulm_score,
           y = mlm_score,
           colour = mean_score)) +
  
    # --- quadrant shading ---
  annotate("rect",
           xmin = -Inf, xmax = 0,
           ymin = -Inf, ymax = 0,
           fill = "#3B6FB6", alpha = 0.08) +

  annotate("rect",
           xmin = 0, xmax = Inf,
           ymin = 0, ymax = Inf,
           fill = "#C0392B", alpha = 0.08) +

  ################
  
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "grey80") +

  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey80") +

  geom_abline(intercept = 0,
              slope = 1,
              colour = "grey60",
              linewidth = 0.5) +

  geom_point(
    aes(size = n_targets,
        colour = neglog10p),
    alpha = 0.85
  ) +

  geom_text_repel(
  aes(label = label),
  fontface = "bold",
  size = 4,
  box.padding = 0.5,
  point.padding = 0.4,
  segment.color = "grey60",
  segment.size = 0.4,
  max.overlaps = Inf
  ) +

  scale_colour_gradient2(
    low = "blue",
    mid = "grey92",
    high = "red",
    midpoint = 0
  ) +

  scale_size(
    range = c(2,10),
    name = "Detected\ntargets"
  ) +

  labs(
    x = "ULM activity score",
    y = "MLM activity score"
  ) +

  coord_equal() +
  
  #coord_cartesian(xlim = c(-3.2, 3.4), ylim = c(-2.9, 3.4)) +
  
  geom_point(
    data = subset(consensus_df, significant),
    shape = 21,
    stroke = 1.2,
    colour = "black",
    fill = NA,
    aes(size = n_targets)
  ) +

  theme_classic(base_size = 14)

ggsave("TF_RAD52STEM.svg", plot = consensus_rad52stem_plot, width = 8, height = 6)
```

The following code checks how many targets of each TF is covered in our protein dataset.

```{r}
covered_targets_AB <- regulons_AB %>%
  filter(target %in% rownames(mat)) %>%
  count(tf)

covered_targets_ABC <- regulons_ABC %>%
  filter(target %in% rownames(mat)) %>%
  count(tf)
```

## Leading Edge genes (GSEA)

Checking Leading Edge genes in the topTable df:

```{r}
# Function to extract the leading-edge genes of one geneset/pathway in a
# fgsea contrast. It then creates a table with a description for each gene
# in the leading-edge

leading_edge_function <- function(fgsea_result, pathway, contrast, filename){
  # See at which row the pathway is indexed and extract the leading-edge genes
  geneset_index <- which(fgsea_result[,1] == pathway)
  leading_edge <- fgsea_result[[geneset_index,8]]
  
  # Create the annotation (description) of each gene
  annot <- queryMany(
    leading_edge,
    scopes = "symbol",
    fields = c("name", "summary"),
    species = "human")
  
  # Save the annotation as a .tsv table
  annot <- data.frame(annot$query, annot$name, annot$summary)
  write.table(x = annot, file = filename, quote = F,
            row.names = F, sep = "\t")
  
  # Returns a table with the TopTable results of the leading-edge genes, so that
  # we can easily check the p-values of those genes
  return(contrast[leading_edge,])
}

leading_edge_function(fgsea_result = fgseaRes.shNT_STEM.shNT_DIFF,
                      pathway = "HALLMARK_DNA_REPAIR",
                      contrast = shNT_STEM.shNT_DIFF,
                      filename = "leading_edge_STEMDIFF_DNA_REPAIR.tsv")
```

RAD52 vs NT, DIFF cells

```{r}
leading_edge_function(fgsea_result = fgseaRes.shRAD52_DIFF.shNT_DIFF,
                      pathway = "HALLMARK_HYPOXIA",
                      contrast = shRAD52_DIFF.shNT_DIFF,
                      filename = "leading_edge_RAD52DIFF_HYPOXIA.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_DIFF.shNT_DIFF,
                      pathway = "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
                      contrast = shRAD52_DIFF.shNT_DIFF,
                      filename = "leading_edge_RAD52DIFF_EMT.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_DIFF.shNT_DIFF,
                      pathway = "HALLMARK_ANGIOGENESIS",
                      contrast = shRAD52_DIFF.shNT_DIFF,
                      filename = "leading_edge_RAD52DIFF_ANGIOGENESIS.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_DIFF.shNT_DIFF,
                      pathway = "HALLMARK_UV_RESPONSE_UP",
                      contrast = shRAD52_DIFF.shNT_DIFF,
                      filename = "leading_edge_RAD52DIFF_UV_UP.tsv")

## fry() identifies some pathways that fgesa() doesn't. Let's look at the genes
## of those fry()-exclusive pathways

p53_genes_RAD52DIFF <- na.omit(shRAD52_DIFF.shNT_DIFF[gene_sets[["HALLMARK_P53_PATHWAY"]],])
p53_genes_RAD52DIFF[order(p53_genes_RAD52DIFF$P.Value),]
```

RAD52 vs NT, STEM cells

```{r}
leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_EMT.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_ESTROGEN_RESPONSE_LATE",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_ESTROGEN.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_COAGULATION",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_COAGULATION.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_APOPTOSIS",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_APOPTOSIS.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_COMPLEMENT",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_COMPLEMENT.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_INFLAMMATORY_RESPONSE",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_INFLAMMATORY.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_PI3K_AKT_MTOR_SIGNALING",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_MTOR.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_UV_RESPONSE_UP",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_UV_UP.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_KRAS_SIGNALING_UP",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_KRAS.tsv")
```

### Heatmaps

Heatmap of protein expression for the GSEA results

```{r}
heatmap_gsea <- function(pathway, protein_data, samples, image_name){
  # Careful: requieres gene_sets to be defined before running the function.
  # Gets the list of genes from the pathway and extracts the expression values
  # from the protein data df
  geneset_prots <- gene_sets[[pathway]]
  # The input argument "samples" has to be a vector specifying the indexes of
  # the samples we are mapping in the protein data df
  protein_matrix <- na.omit(protein_data[geneset_prots, samples])
  
  # Draws and saves the heatmap as svg and jpg
  heatmap <- pheatmap::pheatmap(mat = protein_matrix,
                   scale = 'row', # To scale each protein individually, in case
                                  # one has much higher expression than the rest
                   color = colors.use,
                   border_color = "white")
  
  ggplot2::ggsave(paste0(image_name,".svg"),
                heatmap$gtable,
                width = 8,
                height = 20)

  ggplot2::ggsave(paste0(image_name,".jpg"),
                heatmap$gtable,
                width = 8,
                height = 20,
                dpi = 300)
}

# The same function, but only representing the leading edge genes of the pathway

heatmap_gsea_leading <- function(pathway, protein_data, samples, image_name, fgsea_result){
  # Careful: requieres gene_sets to be defined before running the function.
  # Gets the list of genes from the pathway and extracts the expression values
  # from the protein data df
  geneset_index <- which(fgsea_result[,1] == pathway)
  leading_edge <- fgsea_result[[geneset_index,8]]

  # The input argument "samples" has to be a vector specifying the indexes of
  # the samples we are mapping in the protein data df
  protein_matrix <- na.omit(protein_data[leading_edge, samples])
  
  # Draws and saves the heatmap as svg and jpg
  heatmap <- pheatmap::pheatmap(mat = protein_matrix,
                   scale = 'row', # To scale each protein individually, in case
                                  # one has much higher expression than the rest
                   color = colors.use,
                   border_color = "white")
  
  ggplot2::ggsave(paste0(image_name,".svg"),
                heatmap$gtable,
                width = 8,
                height = 20)

  ggplot2::ggsave(paste0(image_name,".jpg"),
                heatmap$gtable,
                width = 8,
                height = 20,
                dpi = 300)
}

heatmap_gsea_leading(pathway = "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_EMT_leading",
             fgsea_result = fgseaRes.shNT_STEM.shNT_DIFF)

heatmap_gsea_leading(pathway = "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_OXPHOS_leading",
             fgsea_result = fgseaRes.shNT_STEM.shNT_DIFF)

heatmap_gsea_leading(pathway = "HALLMARK_DNA_REPAIR",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_DNAREPAIR_leading",
             fgsea_result = fgseaRes.shNT_STEM.shNT_DIFF)
```
