---
title: "TFM_analysis"
author: "Alejandro Pérez Recio"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing and loading the necessary libraries:

```{r}
# install.packages('readxl')
# BiocManager::install("GSVA")
# install.packages('msigdbr')
library(readxl)
library(NormalyzerDE)
library(FactoMineR)
library(factoextra)
library(limma)

library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)

library(GSVA)
library(msigdbr)
library(GSEABase)
library(tidyverse)

library(fgsea)

library(decoupleR)
library(dorothea)
library(dplyr)
library(tibble)
library(ggplot2)

library(mygene)

library(EnhancedVolcano)
library(ggrepel)
```

Loading the dataset:

```{r}
prots_df <- read_excel("20260413_112142_Glioblastoma y RAD52_Report.xlsx",
                       na = c("", "NaN", "NA"))
head(prots_df)
```

Protein groups are aggregated at the gene level by averaging intensities for proteins mapping to the same gene, to enable compatibility with gene set–based pathway analysis.

We will also separate the samples into two different dataframes to normalyze them separately (stem-like and differentiated) because of the big proteomic differences between these two cell lines.

We will filter out those proteins with a high proportion of NA values. However, we should be careful with the filter, since NA values often derive from low protein quantity, and might be biologically relevant (for example, if the NA values for one protein are mostly in one condition, it is most probably due to that condition having low quantity of that protein, which is informative and relevant, and should not be simply discarded).

Next, we will impute data using Limit of Detection (LoD), using half of the minimum value per protein.

We use a boxplot to represent the global distribution of gene expression in the different samples and check for the presence of noise.

```{r}
# Since we will be working with the PG.Genes IDs, and some proteins in the DF
# share the same PG.Gene (they are isoforms, splice variants, etc.) we will
# collapse the proteins belonging to the same gene into one, using the mean
# expression

df_collapsed <- prots_df %>%
  dplyr::group_by(PG.Genes) %>%
  dplyr::summarise(across(where(is.numeric), mean))

protein_ids <- df_collapsed$PG.Genes

# The original DF contains other data and the replicates are not in adjacent columns.
prots_df_ordered <- data.frame(df_collapsed[2],df_collapsed[6],df_collapsed[10],df_collapsed[14],
                               df_collapsed[3],df_collapsed[7],df_collapsed[11],df_collapsed[15],
                               df_collapsed[4],df_collapsed[8],df_collapsed[12],df_collapsed[16],
                               df_collapsed[5],df_collapsed[9],df_collapsed[13],df_collapsed[17]
                               )

# One of the proteins (Q6ZSR9) has no PG.Gene associated with it (It is encoded
# in the 3'-UTR of AAK1 but there is evidence for the existence of the protein
# from a number of proteomics studies). We will remove this protein for this
# study as we will not be able to map it to a gene. It is located as the last
# row of the data.

prots_df_ordered <- prots_df_ordered[-nrow(prots_df_ordered), ]
protein_ids <- protein_ids[-length(protein_ids)]

rownames(prots_df_ordered) <- protein_ids
colnames(prots_df_ordered) <- c('shNT_DIFF_1','shNT_DIFF_2','shNT_DIFF_3','shNT_DIFF_4',
                                'shNT_STEM_1','shNT_STEM_2','shNT_STEM_3','shNT_STEM_4',
                                'shRAD52_DIFF_1','shRAD52_DIFF_2','shRAD52_DIFF_3','shRAD52_DIFF_4',
                                'shRAD52_STEM_1','shRAD52_STEM_2','shRAD52_STEM_3','shRAD52_STEM_4')

# Separate into 2 different dataframes for the two cell tzpes
prots_DIFF <- data.frame(prots_df_ordered[1],prots_df_ordered[2],prots_df_ordered[3],prots_df_ordered[4],
                         prots_df_ordered[9],prots_df_ordered[10],prots_df_ordered[11],prots_df_ordered[12])

prots_STEM <- data.frame(prots_df_ordered[5],prots_df_ordered[6],prots_df_ordered[7],prots_df_ordered[8],
                         prots_df_ordered[13],prots_df_ordered[14],prots_df_ordered[15],prots_df_ordered[16])

# Filter to keep only those proteins with at least half of non-NA values
prots_DIFF <- prots_DIFF[which(rowMeans(!is.na(prots_DIFF)) > 0.7), ]
prots_STEM <- prots_STEM[which(rowMeans(!is.na(prots_STEM)) > 0.7), ]
prots_all <- prots_df_ordered[which(rowMeans(!is.na(prots_STEM)) > 0.7), ]

# Create protein IDs vectors again with the proteins filtered in each group
protein_ids_DIFF <- rownames(prots_DIFF)
protein_ids_STEM <- rownames(prots_STEM)
protein_ids_all <- rownames(prots_all)

# Quick function to impute NAs with LoD/5

lod5_impute <- function(mat) {
  
  imputed_mat <- t(apply(mat, 1, function(row) {
   
     # Lowest detected value for this protein
    lod <- min(row, na.rm = TRUE)
    
    # Replace NAs with LoD/5
    row[is.na(row)] <- lod / 5
    
    return(row)
  }))
  
  # Preserve row/column names
  rownames(imputed_mat) <- rownames(mat)
  colnames(imputed_mat) <- colnames(mat)
  
  return(imputed_mat)
}

prots_DIFF <- as.data.frame(lod5_impute(prots_DIFF))
prots_STEM <- as.data.frame(lod5_impute(prots_STEM))
prots_all <- as.data.frame(lod5_impute(prots_all))

boxplot(prots_DIFF,outline=F,las=2,ylab="Protein Quantification (Differentiated)")
boxplot(prots_STEM,outline=F,las=2,ylab="Protein Quantification (Stem-like")
```

Variations between replicates of the same conditions show the presence of noise. We will use NormalyzerDE to choose the best normalization technique.

```{r}
## A data frame needs to be generated to specify the experimental design.
design_DIFF <- data.frame(sample=colnames(prots_DIFF),
                     group=c(rep("shNT_DIFF",4),rep("shRAD52_DIFF",4)))

design_STEM <- data.frame(sample=colnames(prots_STEM),
                     group=c(rep("shNT_STEM",4),rep("shRAD52_STEM",4)))

design_all <- data.frame(sample=colnames(prots_df_ordered),
                     group=c(rep("shNT_DIFF",4),
                             rep("shNT_STEM",4),
                             rep("shRAD52_DIFF",4),
                             rep("shRAD52_STEM",4)))

## The experimental design needs to be saved to a file.
write.table(x = design_DIFF, file = "normalyzer_design_DIFF.tsv",quote = F,
            row.names = F, sep = "\t")
write.table(x = design_STEM, file = "normalyzer_design_STEM.tsv",quote = F,
            row.names = F, sep = "\t")
write.table(x = design_all, file = "normalyzer_design_all.tsv",quote = F,
            row.names = F, sep = "\t")

# A table with the raw data is also saved. The values are increased +1 to make
# sure that all values are above 1, so that we can apply log transformations.
prots_DIFF_1 <- prots_DIFF + 1
write.table(x = prots_DIFF_1, file = "prots_DIFF.tsv", quote = F,
            row.names = F, sep = "\t")

prots_STEM_1 <- prots_STEM + 1
write.table(x = prots_STEM_1, file = "prots_STEM.tsv", quote = F,
            row.names = F, sep = "\t")

prots_all_1 <- prots_all + 1
write.table(x = prots_all_1, file = "prots_all.tsv", quote = F,
            row.names = F, sep = "\t")

## The following function runs all normalization techniques and generates
## evaluations measures for them. 
normalyzer(jobName = "prots_diff",designPath = "normalyzer_design_DIFF.tsv",
           dataPath = "prots_DIFF.tsv",outputDir = ".")

normalyzer(jobName = "prots_stem",designPath = "normalyzer_design_STEM.tsv",
           dataPath = "prots_STEM.tsv",outputDir = ".")

normalyzer(jobName = "prots_all",designPath = "normalyzer_design_all.tsv",
           dataPath = "prots_all.tsv",outputDir = ".")
```

We will pick Variance Stabilizing Normalization (VSN), as it seems to be reducing intragroup variance the best.

```{r}
normalized_prot_DIFF <- read.table(file="prots_diff/VSN-normalized.txt", header=T)
rownames(normalized_prot_DIFF) <- protein_ids_DIFF
boxplot(normalized_prot_DIFF,outline=F,las=2,ylab="Normalized Protein Quantification (Differentiated)")

normalized_prot_STEM <- read.table(file="prots_stem/VSN-normalized.txt", header=T)
rownames(normalized_prot_STEM) <- protein_ids_STEM
boxplot(normalized_prot_STEM,outline=F,las=2,ylab="Normalized Protein Quantification (Stem-like)")

normalized_protein <- read.table(file="prots_all/VSN-normalized.txt", header=T)
rownames(normalized_protein) <- protein_ids_all
boxplot(normalized_protein,outline=F,las=2,ylab="Normalized Protein Quantification (Stem-like)")
```

We can represent two different replicates in a scatter plot to check that they are mostly similar:

```{r}
plot(x = normalized_prot_DIFF$shNT_DIFF_1,
     y = normalized_prot_DIFF$shNT_DIFF_2,
     pch=19,col="grey",xlab="shNT_DIFF_1",ylab="shNT_DIFF_2",cex=0.5)
text(x=4,y=12,
     labels = paste(c(
      "cor = ",
      round(100*cor(normalized_prot_DIFF$shNT_DIFF_1,
                    normalized_prot_DIFF$shNT_DIFF_2),
            digits = 2),
      "%"), collapse=""))
```

We will analyze the data with Principal Component Analysis (PCA) in order to verify that no mistakes were made in labeling the samples and that samples with similar biological conditions cluster together as expected.

```{r}
# TransSTEMing the df for the function to work
pca_prot_diff <- data.frame(colnames(normalized_prot_DIFF),
                                  t(normalized_prot_DIFF))
colnames(pca_prot_diff)[1] <- "Sample"
pca_prot_diff[1:8][1:6]

res_pca_diff <- PCA(pca_prot_diff, 
               graph = FALSE,
               scale.unit = TRUE,
               quali.sup = 1 )

fviz_pca_ind(res_pca_diff, col.ind = design_DIFF$group, 
             pointsize=2, pointshape=21,fill="black",
             repel = TRUE, 
             addEllipses = TRUE,ellipse.type = "confidence",
             legend.title="Conditions",
             title="",
             show_legend=TRUE,show_guide=TRUE)


pca_prot_stem <- data.frame(colnames(normalized_prot_STEM),
                                  t(normalized_prot_STEM))
colnames(pca_prot_stem)[1] <- "Sample"
pca_prot_stem[1:8][1:6]

res_pca_stem <- PCA(pca_prot_stem, 
               graph = FALSE,
               scale.unit = TRUE,
               quali.sup = 1 )

fviz_pca_ind(res_pca_stem, col.ind = design_STEM$group, 
             pointsize=2, pointshape=21,fill="black",
             repel = TRUE, 
             addEllipses = TRUE,ellipse.type = "confidence",
             legend.title="Conditions",
             title="",
             show_legend=TRUE,show_guide=TRUE)
```

The differences between cells with the Non-target short hairpin interference RNA (shNT) and the ones with target interference RNA against RAD52 (shRAD52) are clearly seen in DIFF (Glioblastoma Differentiated Cells), but not so much for the STEM (Glioblastoma Stem Cells).

We can also perform hierarchical clustering analysis over the PCA results and represent it in a dendrogram to check that the different conditions group up together as expected.

```{r}
res_hcpc_diff <- HCPC(res_pca_diff, graph=FALSE, nb.clust = 4)   

fviz_dend(res_hcpc_diff,k=4,
          cex = 0.75,                       # Label size
          palette = "jco",               # Color palette see ?ggpubr::ggpar
          rect = TRUE, rect_fill = TRUE, # Add rectangle around groups
          rect_border = "jco",           # Rectangle color
          type="rectangle",
          labels_track_height = 700      # Augment the room for labels
)

res_pca_stem <- HCPC(res_pca_stem, graph=FALSE,nb.clust = 4)   

fviz_dend(res_pca_stem,k=4,
          cex = 0.75,                       # Label size
          palette = "jco",               # Color palette see ?ggpubr::ggpar
          rect = TRUE, rect_fill = TRUE, # Add rectangle around groups
          rect_border = "jco",           # Rectangle color
          type="rectangle",
          labels_track_height = 700      # Augment the room for labels
)
```

As we saw in the PCA, the sampls for STEM cells are very similar and the hierarchical clustering cannot differentiate between those with and without shRAD52. In the DIFF cells, the clustering differentiates better between the two conditions, but still mixes up some of the samples.

## DEGs

Differential expression analysis with limma:

### Individual limma contrasts

EXPERIMENTAL: Contrastes individualmente, por separado en diferentes diseños, y usando pvalor en lugar de ajustado:

```{r}
intersect_proteins <- intersect(rownames(normalized_prot_DIFF), rownames(normalized_prot_STEM))

experimental_design_diff <- model.matrix(~ -1+factor(c(1,1,1,1,2,2,2,2)))
colnames(experimental_design_diff) <- c("shNT_DIFF","shRAD52_DIFF")
experimental_design_stem <- model.matrix(~ -1+factor(c(1,1,1,1,2,2,2,2)))
colnames(experimental_design_stem) <- c("shNT_STEM","shRAD52_STEM")

normalized_prot_DIFF <- normalized_prot_DIFF[intersect_proteins,]

## Linear model fit to the experimental design, mean expression calculation
linear_fit_diff <- lmFit(normalized_prot_DIFF, experimental_design_diff)
linear_fit_stem <- lmFit(normalized_prot_STEM, experimental_design_stem)

## Contrast specification treatment-control
contrast_matrix_diff <- makeContrasts(shRAD52_DIFF-shNT_DIFF,
                                 levels=c("shNT_DIFF","shRAD52_DIFF"))
contrast_matrix_stem <- makeContrasts(shRAD52_STEM-shNT_STEM,
                                 levels=c("shNT_STEM","shRAD52_STEM"))

contrast_linear_fit_diff <- contrasts.fit(linear_fit_diff, contrast_matrix_diff)
contrast_results_diff <- eBayes(contrast_linear_fit_diff)
contrast_linear_fit_stem <- contrasts.fit(linear_fit_stem, contrast_matrix_stem)
contrast_results_stem <- eBayes(contrast_linear_fit_stem)
```

```{r}
## Results extraction for the first contrast (shRAD52_DIFF.shNT_DIFF) sorted according to logFC
shRAD52_DIFF.shNT_DIFF <- topTable(contrast_results_diff, 
                      number=nrow(normalized_prot_DIFF),
                      coef=1,
                      sort.by="logFC")
head(shRAD52_DIFF.shNT_DIFF)

## DEGs extraction
limma.log2fc <- shRAD52_DIFF.shNT_DIFF$logFC
limma.adjusted.pval <- shRAD52_DIFF.shNT_DIFF$adj.P.Val 
limma.pval <- shRAD52_DIFF.shNT_DIFF$P.Value 
limma.protein_ids <- rownames(shRAD52_DIFF.shNT_DIFF)

names(limma.log2fc) <- limma.protein_ids
names(limma.adjusted.pval) <- limma.protein_ids
names(limma.pval) <- limma.protein_ids

shRAD52_DIFF.shNT_DIFF.activated.genes <- limma.protein_ids[limma.log2fc > 0 & 
                                          limma.pval < 0.01]
shRAD52_DIFF.shNT_DIFF.repressed.genes <- limma.protein_ids[limma.log2fc < 0 & 
                                          limma.pval < 0.01]
length(shRAD52_DIFF.shNT_DIFF.activated.genes)
length(shRAD52_DIFF.shNT_DIFF.repressed.genes)

## Volcano plot
volcano_df <- data.frame(
  logFC = limma.log2fc,
  pval = limma.pval
)

volcano_df$negLogP <- -log10(volcano_df$pval)
volcano_df$gene <- rownames(volcano_df)

volcano_df$group <- "NS"
volcano_df$group[volcano_df$gene %in% shRAD52_DIFF.shNT_DIFF.activated.genes] <- "Up"
volcano_df$group[volcano_df$gene %in% shRAD52_DIFF.shNT_DIFF.repressed.genes] <- "Down"

# pick top genes for labeling (by significance)
top_genes <- volcano_df[order(volcano_df$pval), ][1:10, ]
max_fc <- max(abs(volcano_df$logFC), na.rm = TRUE) * 1.05

volcano <- ggplot(volcano_df, aes(x = logFC, y = negLogP)) +

  # background points (neutral grey, transparent)
  geom_point(
    color = "grey75",
    alpha = 0.6,
    size = 1.3
  ) +

  # highlighted groups
  geom_point(
    data = subset(volcano_df, group == "Up"),
    color = "#2C7BB6",
    alpha = 0.85,
    size = 1.5
  ) +
  geom_point(
    data = subset(volcano_df, group == "Down"),
    color = "#D7191C",
    alpha = 0.85,
    size = 1.5
  ) +

  # threshold lines (subtle, not distracting)
  geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = "grey60") +

  # label only top genes
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey50"
  ) +

  # axis + theme styling (this is what makes it “Nature-like”)
  labs(
    x = "log2 fold change",
    y = expression(-log[10]("p-value"))
  ) +
  
  coord_cartesian(xlim = c(-max_fc, max_fc),
                  ylim = c(0,6.5)) +

  theme_classic(base_size = 14) +
  theme(

    # remove minor clutter
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),

    # cleaner axis text
    axis.text = element_text(color = "black"),

    # no grid anywhere (very important for journal look)
    panel.grid = element_blank(),

    # slightly stronger axis titles
    axis.title = element_text(face = "bold"),

    # tight margins (optional but very “paper-like”)
    plot.margin = margin(10, 10, 10, 10)
  )

ggsave("volcano_RAD52DIFF.svg", plot = volcano, width = 6, height = 6)

################################ STEM RAD52 vs NT #########################

## Results extraction for the second contrast (shRAD52_STEM.shNT_STEM) sorted according to logFC
shRAD52_STEM.shNT_STEM <- topTable(contrast_results_stem, 
                      number=nrow(normalized_prot_STEM),
                      coef=1,
                      sort.by="logFC")
head(shRAD52_STEM.shNT_STEM)

## DEGs extraction
limma.log2fc <- shRAD52_STEM.shNT_STEM$logFC
limma.adjusted.pval <- shRAD52_STEM.shNT_STEM$adj.P.Val 
limma.pval <- shRAD52_STEM.shNT_STEM$P.Value 
limma.protein_ids <- rownames(shRAD52_STEM.shNT_STEM)

names(limma.log2fc) <- limma.protein_ids
names(limma.adjusted.pval) <- limma.protein_ids
names(limma.pval) <- limma.protein_ids

shRAD52_STEM.shNT_STEM.activated.genes <- limma.protein_ids[limma.log2fc > 0 & 
                                          limma.pval < 0.01]
shRAD52_STEM.shNT_STEM.repressed.genes <- limma.protein_ids[limma.log2fc < 0 & 
                                          limma.pval < 0.01]
length(shRAD52_STEM.shNT_STEM.activated.genes)
length(shRAD52_STEM.shNT_STEM.repressed.genes)

## Volcano plot
volcano_df <- data.frame(
  logFC = limma.log2fc,
  pval = limma.pval
)

volcano_df$negLogP <- -log10(volcano_df$pval)
volcano_df$gene <- rownames(volcano_df)

volcano_df$group <- "NS"
volcano_df$group[volcano_df$gene %in% shRAD52_STEM.shNT_STEM.activated.genes] <- "Up"
volcano_df$group[volcano_df$gene %in% shRAD52_STEM.shNT_STEM.repressed.genes] <- "Down"

# pick top genes for labeling (by significance)
top_genes <- volcano_df[order(volcano_df$pval), ][1:10, ]
max_fc <- max(abs(volcano_df$logFC), na.rm = TRUE) * 1.05

volcano <- ggplot(volcano_df, aes(x = logFC, y = negLogP)) +

  # background points (neutral grey, transparent)
  geom_point(
    color = "grey75",
    alpha = 0.6,
    size = 1.3
  ) +

  # highlighted groups
  geom_point(
    data = subset(volcano_df, group == "Up"),
    color = "#2C7BB6",
    alpha = 0.85,
    size = 1.5
  ) +
  geom_point(
    data = subset(volcano_df, group == "Down"),
    color = "#D7191C",
    alpha = 0.85,
    size = 1.5
  ) +

  # threshold lines (subtle, not distracting)
  geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = "grey60") +

  # label only top genes
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey50"
  ) +

  # axis + theme styling (this is what makes it “Nature-like”)
  labs(
    x = "log2 fold change",
    y = expression(-log[10]("p-value"))
  ) +
  
  coord_cartesian(xlim = c(-max_fc, max_fc)) +

  theme_classic(base_size = 14) +
  theme(

    # remove minor clutter
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),

    # cleaner axis text
    axis.text = element_text(color = "black"),

    # no grid anywhere (very important for journal look)
    panel.grid = element_blank(),

    # slightly stronger axis titles
    axis.title = element_text(face = "bold"),

    # tight margins (optional but very “paper-like”)
    plot.margin = margin(10, 10, 10, 10)
  )

ggsave("volcano_RAD52STEM.svg", plot = volcano, width = 6, height = 6)
```

### Grouped limma analysis

```{r}
## Experimental design specification

## Experimental design specification
experimental.design <- model.matrix(~ -1+factor(c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,4)))
colnames(experimental.design) <- c("shNT_DIFF","shNT_STEM","shRAD52_DIFF","shRAD52_STEM")

## Linear model fit to the experimental design, mean expression calculation
linear.fit <- lmFit(normalized_protein, experimental.design)

## Contrast specification treatment-control
contrast.matrix <- makeContrasts(shRAD52_DIFF-shNT_DIFF,shRAD52_STEM-shNT_STEM,shRAD52_STEM-shRAD52_DIFF,shNT_STEM-shNT_DIFF,
                                 levels=c("shNT_DIFF","shNT_STEM","shRAD52_DIFF","shRAD52_STEM"))

contrast.linear.fit <- contrasts.fit(linear.fit, contrast.matrix)
contrast.results <- eBayes(contrast.linear.fit, robust = TRUE)
# Trying with robust=TRUE to try and make it less sensible to variance imbalance

```

Fourth contrast: shNT_STEM vs shNT_DIFF.

```{r}
## Results extraction for the fourth contrast (coef=4) sorted according to logFC
shNT_STEM.shNT_DIFF <- topTable(contrast.results, 
                      number=nrow(normalized_protein),
                      coef=4,
                      sort.by="logFC")
head(shNT_STEM.shNT_DIFF)

## DEGs extraction
limma.log2fc <- shNT_STEM.shNT_DIFF$logFC
limma.adjusted.pval <- shNT_STEM.shNT_DIFF$adj.P.Val
limma.pval <- shNT_STEM.shNT_DIFF$P.Value 
limma.protein_ids <- rownames(shNT_STEM.shNT_DIFF)

names(limma.log2fc) <- limma.protein_ids
names(limma.adjusted.pval) <- limma.protein_ids
names(limma.pval) <- limma.protein_ids

shNT_STEM.shNT_DIFF.activated.genes <- limma.protein_ids[limma.log2fc > 0.58 & 
                                          limma.adjusted.pval < 0.01]
shNT_STEM.shNT_DIFF.repressed.genes <- limma.protein_ids[limma.log2fc < - 0.58 & 
                                          limma.adjusted.pval < 0.01]
length(shNT_STEM.shNT_DIFF.activated.genes)
length(shNT_STEM.shNT_DIFF.repressed.genes)

## Volcano plot
volcano_df <- data.frame(
  logFC = limma.log2fc,
  padj = limma.adjusted.pval
)

volcano_df$negLogP <- -log10(volcano_df$padj)
volcano_df$gene <- rownames(volcano_df)

volcano_df$group <- "NS"
volcano_df$group[volcano_df$gene %in% shNT_STEM.shNT_DIFF.activated.genes] <- "Up"
volcano_df$group[volcano_df$gene %in% shNT_STEM.shNT_DIFF.repressed.genes] <- "Down"

# pick top genes for labeling (by significance)
top_genes <- volcano_df[order(volcano_df$padj), ][1:10, ]
max_fc <- max(abs(volcano_df$logFC), na.rm = TRUE) * 1.05

volcano <- ggplot(volcano_df, aes(x = logFC, y = negLogP)) +

  # background points (neutral grey, transparent)
  geom_point(
    color = "grey75",
    alpha = 0.6,
    size = 1.3
  ) +

  # highlighted groups
  geom_point(
    data = subset(volcano_df, group == "Up"),
    color = "#2C7BB6",
    alpha = 0.85,
    size = 1.5
  ) +
  geom_point(
    data = subset(volcano_df, group == "Down"),
    color = "#D7191C",
    alpha = 0.85,
    size = 1.5
  ) +

  # threshold lines (subtle, not distracting)
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey60") +
  geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = "grey60") +

  # label only top genes
  geom_text_repel(
    data = top_genes,
    aes(label = gene),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey50"
  ) +

  # axis + theme styling (this is what makes it “Nature-like”)
  labs(
    x = "log2 fold change",
    y = expression(-log[10]("adjusted p-value"))
  ) +
  
  coord_cartesian(xlim = c(-max_fc, max_fc)) +

  theme_classic(base_size = 14) +
  theme(

    # remove minor clutter
    axis.line = element_line(color = "black", linewidth = 0.6),
    axis.ticks = element_line(color = "black"),

    # cleaner axis text
    axis.text = element_text(color = "black"),

    # no grid anywhere (very important for journal look)
    panel.grid = element_blank(),

    # slightly stronger axis titles
    axis.title = element_text(face = "bold"),

    # tight margins (optional but very “paper-like”)
    plot.margin = margin(10, 10, 10, 10)
  )

ggsave("volcano_STEMDIFF.svg", plot = volcano, width = 6, height = 6)
```

### Interaction test

Essentially answering the question: which proteins change in STEM cells after adding RAD52 that do not change in DIFF cells after RAD52 addition?

```{r}
CellType <- factor(c(
  rep("DIFF", 8),
  rep("STEM", 8)))

Knockdown <- factor(c(
  rep("shNT", 4),
  rep("shRAD52", 4),
  rep("shNT", 4),
  rep("shRAD52", 4)))

design <- model.matrix(~0 + CellType:Knockdown)
colnames(design) <- c(
  "DIFF_shNT",
  "STEM_shNT",
  "DIFF_shRAD52",
  "STEM_shRAD52")

cont <- makeContrasts(
  RAD52_DIFF = DIFF_shRAD52 - DIFF_shNT,
  RAD52_STEM = STEM_shRAD52 - STEM_shNT,
  Interaction =
    (STEM_shRAD52 - STEM_shNT) -
    (DIFF_shRAD52 - DIFF_shNT),
  levels = design)

fit <- lmFit(normalized_protein, design)
fit <- contrasts.fit(fit, cont)
fit <- eBayes(fit)

interaction.contrast <- topTable(fit, 
                      number=nrow(normalized_protein),
                      coef="Interaction",
                      sort.by="logFC")

## DEGs extraction
limma.log2fc <- interaction.contrast$logFC
limma.adjusted.pval <- interaction.contrast$adj.P.Val
limma.pval <- interaction.contrast$P.Value
limma.protein_ids <- rownames(interaction.contrast)

names(limma.log2fc) <- limma.protein_ids
names(limma.adjusted.pval) <- limma.protein_ids
names(limma.pval) <- limma.protein_ids

interaction.contrast.activated.genes <- limma.protein_ids[limma.log2fc > 0.58 & 
                                          limma.pval < 0.01]
interaction.contrast.repressed.genes <- limma.protein_ids[limma.log2fc < - 0.58 & 
                                          limma.pval < 0.01]
length(interaction.contrast.activated.genes)
length(interaction.contrast.repressed.genes)

## Volcano plot
log.limma.adjusted.pval <- -log10(limma.pval)
plot(limma.log2fc,log.limma.adjusted.pval,pch=19,col="grey",cex=0.8,
     xlim=c(-6,6),ylim = c(0,10), 
     xlab="log2(Fold-chage)",ylab="-log10(p-val)",cex.lab=1.5)

points(x = limma.log2fc[interaction.contrast.activated.genes],
       y = log.limma.adjusted.pval[interaction.contrast.activated.genes],col="blue",cex=0.8,pch=19)
points(x = limma.log2fc[interaction.contrast.repressed.genes],
       y = log.limma.adjusted.pval[interaction.contrast.repressed.genes],col="red",cex=0.8,pch=19)
```

## Enrichment

### GO Enrichment

With the first comparison shRAD52_DIFF vs shNT_DIFF (activated and repressed genes). Even with a pvalueCutoff of 0.1, there are almost no enriched terms (there are 2 of them in the repressed genes, related to mithocondrial membrane)

```{r}
activated.enrich.go.shRAD52_DIFF.shNT_DIFF <- enrichGO(gene = shRAD52_DIFF.shNT_DIFF.activated.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.1,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(activated.enrich.go.shRAD52_DIFF.shNT_DIFF)
#treeplot(pairwise_termsim(activated.enrich.go.shRAD52_DIFF.shNT_DIFF),showCategory = 20)

repressed.enrich.go.shRAD52_DIFF.shNT_DIFF <- enrichGO(gene = shRAD52_DIFF.shNT_DIFF.repressed.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.1,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(repressed.enrich.go.shRAD52_DIFF.shNT_DIFF)
#treeplot(pairwise_termsim(repressed.enrich.go.shRAD52_DIFF.shNT_DIFF),showCategory = 20)

```

Now, for shRAD52_STEM vs shNT_STEM

```{r}
activated.enrich.go.shRAD52_STEM.shNT_STEM <- enrichGO(gene = shRAD52_STEM.shNT_STEM.activated.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.1,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(activated.enrich.go.shRAD52_STEM.shNT_STEM)
#treeplot(pairwise_termsim(activated.enrich.go.shRAD52_DIFF.shNT_DIFF),showCategory = 20)

repressed.enrich.go.shRAD52_STEM.shNT_STEM <- enrichGO(gene = shRAD52_STEM.shNT_STEM.repressed.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.1,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(repressed.enrich.go.shRAD52_STEM.shNT_STEM)
#treeplot(pairwise_termsim(repressed.enrich.go.shRAD52_DIFF.shNT_DIFF),showCategory = 20)
```

And again with shNT_STEM vs shNT_DIFF:

```{r}
activated.enrich.go.shNT_STEM.shNT_DIFF <- enrichGO(gene = shNT_STEM.shNT_DIFF.activated.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.5,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(activated.enrich.go.shNT_STEM.shNT_DIFF)
#treeplot(pairwise_termsim(activated.enrich.go.shNT_STEM.shNT_DIFF),showCategory = 20)
#barplot(activated.enrich.go.shNT_STEM.shNT_DIFF, showCategory = 20)

repressed.enrich.go.shNT_STEM.shNT_DIFF <- enrichGO(gene = shNT_STEM.shNT_DIFF.repressed.genes,
                                OrgDb         = org.Hs.eg.db,
                                ont           = "BP",
                                pAdjustMethod = "BH",
                                pvalueCutoff  = 0.5,
                                readable      = FALSE,
                                keyType = "SYMBOL")

dotplot(repressed.enrich.go.shNT_STEM.shNT_DIFF)
treeplot(pairwise_termsim(repressed.enrich.go.shNT_STEM.shNT_DIFF),showCategory = 20)
barplot(repressed.enrich.go.shNT_STEM.shNT_DIFF, showCategory = 20)
```

### KEGGs

Now, we will try to obtain enriched KEGG pathways for the shRAD52_DIFF-shNT_DIFF comparison:

```{r}
activated.df <- bitr(
    shRAD52_DIFF.shNT_DIFF.activated.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

repressed.df <- bitr(
    shRAD52_DIFF.shNT_DIFF.repressed.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

universe.df <- bitr(
    protein_ids,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

activated.enrich.kegg.shRAD52_DIFF.shNT_DIFF <- enrichKEGG(
    gene = activated.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- activated.enrich.kegg.shRAD52_DIFF.shNT_DIFF@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

repressed.enrich.kegg.shRAD52_DIFF.shNT_DIFF <- enrichKEGG(
    gene = repressed.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- repressed.enrich.kegg.shRAD52_DIFF.shNT_DIFF@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)
```

For the shRAD52_STEM-shNT_STEM comparison:

```{r}
activated.df <- bitr(
    shRAD52_STEM.shNT_STEM.activated.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

repressed.df <- bitr(
    shRAD52_STEM.shNT_STEM.repressed.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

universe.df <- bitr(
    protein_ids,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

activated.enrich.kegg.shRAD52_STEM.shNT_STEM <- enrichKEGG(
    gene = activated.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- activated.enrich.kegg.shRAD52_STEM.shNT_STEM@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

repressed.enrich.kegg.shRAD52_STEM.shNT_STEM <- enrichKEGG(
    gene = repressed.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- repressed.enrich.kegg.shRAD52_STEM.shNT_STEM@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

```

For the record, the shNT_STEM-shNT_DIFF comparison:

```{r}
activated.df <- bitr(
    shNT_STEM.shNT_DIFF.activated.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

repressed.df <- bitr(
    shNT_STEM.shNT_DIFF.repressed.genes,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

universe.df <- bitr(
    protein_ids,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db)

activated.enrich.kegg.shNT_STEM.shNT_DIFF <- enrichKEGG(
    gene = activated.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- activated.enrich.kegg.shNT_STEM.shNT_DIFF@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

repressed.enrich.kegg.shNT_STEM.shNT_DIFF <- enrichKEGG(
    gene = repressed.df$ENTREZID,
    universe = universe.df$ENTREZID,
    organism = "hsa",
    pvalueCutoff  = 1)

res <- repressed.enrich.kegg.shNT_STEM.shNT_DIFF@result

res <- res[order(res$pvalue), ]
head(res[, c("Description", "Count", "pvalue", "p.adjust")], 15)

```

## GSEA (Gene Set Enrichment Analysis)

### fgsea()

Another way of obtaining enriched pathways. We will first try with fgsea(). First, the differentiated cells shRAD52_DIFF vs shNT_DIFF:

```{r}
set.seed(123) # Setting a fixed seed, because GSEA calculates the pvalues with
              # Monte-Carlo sampling and so they may vary

msig <- msigdbr(species = "Homo sapiens", collection = "H")  # Hallmark sets

gene_sets <- msig %>%
  dplyr::select(gs_name, gene_symbol) %>%
  split(x = .$gene_symbol, f = .$gs_name)

ranks <- shRAD52_DIFF.shNT_DIFF$t
names(ranks) <- rownames(shRAD52_DIFF.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

fgseaRes.shRAD52_DIFF.shNT_DIFF <- fgsea(
  pathways = gene_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)
```

There seems to be some pathways weakly enriched (only pvalue is <0.05 for some, but no adjusted pvalue). Let's do the other comparisons:

```{r}
ranks <- shRAD52_STEM.shNT_STEM$t
names(ranks) <- rownames(shRAD52_STEM.shNT_STEM)

ranks <- sort(ranks, decreasing = TRUE)

fgseaRes.shRAD52_STEM.shNT_STEM <- fgsea(
  pathways = gene_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

###################

ranks <- shRAD52_STEM.shRAD52_DIFF$t
names(ranks) <- rownames(shRAD52_STEM.shRAD52_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

fgseaRes.shRAD52_STEM.shRAD52_DIFF <- fgsea(
  pathways = gene_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

###################

ranks <- shNT_STEM.shNT_DIFF$t
names(ranks) <- rownames(shNT_STEM.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

fgseaRes.shNT_STEM.shNT_DIFF <- fgsea(
  pathways = gene_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)
```

#### FGSEA Plots

For shNT_STEM.shNT_DIFF:

```{r}
ranks <- shNT_STEM.shNT_DIFF$t
names(ranks) <- rownames(shNT_STEM.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

plotEnrichment(gene_sets[["HALLMARK_DNA_REPAIR"]], ranks)

###### NES Barplot

plot_df <- fgseaRes.shNT_STEM.shNT_DIFF %>%
  filter(padj < 0.05) %>%
  arrange(NES)

NES_barplot <- ggplot(plot_df,
               aes(y = reorder(pathway, NES),
                   x = NES,
                   fill = NES)) +
  geom_col() +
  geom_vline(xintercept = 0,
             linetype = "dashed") +
  theme_classic()

ggsave("NES_Barplot_STEMDIFF.svg", plot = NES_barplot, width = 9, height = 6)

###### Dot plot

dotplot <- ggplot(plot_df,
       aes(NES, reorder(pathway, NES))) +
    geom_point(aes(size = size,
                   color = padj))

ggsave("Dotplot_STEMDIFF.svg", plot = dotplot, width = 8, height = 6)
```

For shRAD52_STEM.shNT_STEM and shRAD52_DIFF.shNT_DIFF:

```{r}
plot_df <- fgseaRes.shRAD52_STEM.shNT_STEM %>%
  filter(pval < 0.05) %>%
  arrange(NES)

dotplot <- ggplot(plot_df,
       aes(NES, reorder(pathway, NES))) +
    geom_point(aes(size = size,
                   color = pval))

ggsave("Dotplot_RAD52STEM.svg", plot = dotplot, width = 8, height = 6)

################

plot_df <- fgseaRes.shRAD52_DIFF.shNT_DIFF %>%
  filter(pval < 0.05) %>%
  arrange(NES)

dotplot <- ggplot(plot_df,
       aes(NES, reorder(pathway, NES))) +
    geom_point(aes(size = size,
                   color = pval))

ggsave("Dotplot_RAD52DIFF.svg", plot = dotplot, width = 8, height = 6)

```

And for the Interaction comparison, previously defined:

```{r}
ranks <- interaction.contrast$t
names(ranks) <- rownames(interaction.contrast)

ranks <- sort(ranks, decreasing = TRUE)

fgseaRes.interaction.contrast <- fgsea(
  pathways = gene_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)
```

Not much to extract from here. Adjusted p-values are all high, and the only p.value < 0.05 is HALLMARK_UV_RESPONSE_UP, which would mean the reduction of the expression of RAD52 with the shRNA has different effects ond STEM and DIFF cells, and in STEM cells there is a reduction of the UV Response that is not seen in DIFF cells. Considering that UV Response might include genes related with DNA reparation and damage, it could be of interest.

### camera()

Now I will try with camera(), part of the limma package, with the shRAD52_DIFF-shNT_DIFF comparison as an example:

```{r}
cameraRes.shRAD52_DIFF.shNT_DIFF <- camera(
    y = normalized_prot_DIFF,
    index = gene_sets,
    design = experimental_design_diff,
    contrast = contrast_matrix_diff[,1])

cameraRes.shRAD52_STEM.shNT_STEM <- camera(
    y = normalized_prot_STEM,
    index = gene_sets,
    design = experimental_design_stem,
    contrast = contrast_matrix_stem[,1])

# cameraRes.interaction.contrast <- camera(
#     y = normalized_protein,
#     index = gene_sets,
#     design = design,
#     contrast = cont[,1])

# No relevance for interaction.contrast

cameraRes.shNT_STEM.shNT_DIFF <- camera(
    y = normalized_protein,
    index = gene_sets,
    design = experimental.design,
    contrast = contrast.matrix[,4])
```

### fry()

It seems to obtain somewhat similar results, the first 5 pathways are the same, although in different order and different pvalues.

Now fry():

```{r}
fryRes.shRAD52_DIFF.shNT_DIFF <- fry(
  normalized_prot_DIFF,
  index = gene_sets,
  design = experimental_design_diff,
  contrast = contrast_matrix_diff[,1])

fryRes.shRAD52_STEM.shNT_STEM <- fry(
  normalized_prot_STEM,
  index = gene_sets,
  design = experimental_design_stem,
  contrast = contrast_matrix_stem[,1])

fryRes.interaction.contrast <- fry(
  normalized_protein,
  index = gene_sets,
  design = design,
  contrast = cont[,1])

fryRes.shNT_STEM.shNT_DIFF <- fry(
  normalized_protein,
  index = gene_sets,
  design = experimental.design,
  contrast = contrast.matrix[,4])
```

## GSVA (Gene Set Variation Analysis)

### Individual analysis

```{r}
msig <- msigdbr(species = "Homo sapiens", collection = "H")  # Hallmark sets

gene_sets <- msig %>%
  dplyr::select(gs_name, gene_symbol) %>%
  split(x = .$gene_symbol, f = .$gs_name)

# shRAD52_DIFF-shNT_DIFF
expr_mat_diff <- as.matrix(normalized_prot_DIFF)

param_diff <- gsvaParam(
  exprData = expr_mat_diff,
  geneSets = gene_sets,
  kcdf = "Gaussian"
)

gsva_res_diff <- gsva(param_diff)

## Linear model fit to the experimental design, mean expression calculation
gsva.linear.fit <- lmFit(gsva_res_diff, experimental_design_diff)

## Contrast specification treatment-control
gsva.contrast.matrix <- makeContrasts(shRAD52_DIFF-shNT_DIFF,
                                 levels=c("shNT_DIFF","shRAD52_DIFF"))

gsva.contrast.linear.fit <- contrasts.fit(gsva.linear.fit, gsva.contrast.matrix)
gsva.contrast.results <- eBayes(gsva.contrast.linear.fit)

## Results extraction for the first contrast (shRAD52_DIFF - shNT_DIFF) sorted according to logFC
gsva.shRAD52_DIFF.shNT_DIFF <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res_diff),
                      coef=1,
                      sort.by="logFC")
head(gsva.shRAD52_DIFF.shNT_DIFF)

##############################################################################

# shRAD52_STEM-shNT_STEM
expr_mat_stem <- as.matrix(normalized_prot_STEM)

param_stem <- gsvaParam(
  exprData = expr_mat_stem,
  geneSets = gene_sets,
  kcdf = "Gaussian"
)

gsva_res_stem <- gsva(param_stem)

## Linear model fit to the experimental design, mean expression calculation
gsva.linear.fit <- lmFit(gsva_res_stem, experimental_design_stem)

## Contrast specification treatment-control
gsva.contrast.matrix <- makeContrasts(shRAD52_STEM-shNT_STEM,
                                 levels=c("shNT_STEM","shRAD52_STEM"))

gsva.contrast.linear.fit <- contrasts.fit(gsva.linear.fit, gsva.contrast.matrix)
gsva.contrast.results <- eBayes(gsva.contrast.linear.fit)

## Results extraction for the second contrast (shRAD52_STEM - shNT_STEM) sorted according to logFC
gsva.shRAD52_STEM.shNT_STEM <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res_stem),
                      coef=1,
                      sort.by="logFC")
head(gsva.shRAD52_STEM.shNT_STEM)
```


### Grouped analysis

All of the contrasts, made in grouped fashion:

```{r}
expr_mat <- as.matrix(normalized_protein)

param <- gsvaParam(
  exprData = expr_mat,
  geneSets = gene_sets,
  kcdf = "Gaussian"
)

gsva_res <- gsva(param)

## Linear model fit to the experimental design, mean expression calculation
gsva.linear.fit <- lmFit(gsva_res, experimental.design)

## Contrast specification treatment-control
gsva.contrast.matrix <- makeContrasts(shRAD52_DIFF-shNT_DIFF,shRAD52_STEM-shNT_STEM,
                                      shRAD52_STEM-shRAD52_DIFF,shNT_STEM-shNT_DIFF,
                                 levels=c("shNT_DIFF","shNT_STEM","shRAD52_DIFF","shRAD52_STEM"))

gsva.contrast.linear.fit <- contrasts.fit(gsva.linear.fit, gsva.contrast.matrix)
gsva.contrast.results <- eBayes(gsva.contrast.linear.fit)

## Results extraction for the first contrast shRAD52_DIFF - shNT_DIFF) sorted according to logFC
grouped_gsva.shRAD52_DIFF.shNT_DIFF <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res),
                      coef=1,
                      sort.by="P")
head(grouped_gsva.shRAD52_DIFF.shNT_DIFF)

## Results extraction for the second contrast (shRAD52_STEM - shNT_STEM) sorted according to logFC
grouped_gsva.shRAD52_STEM.shNT_STEM <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res),
                      coef=2,
                      sort.by="logFC")
head(grouped_gsva.shRAD52_STEM.shNT_STEM)

## Results extraction for the third contrast (shRAD52_STEM - shRAD52_DIFF) sorted according to logFC
grouped_gsva.shRAD52_STEM.shRAD52_DIFF <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res),
                      coef=3,
                      sort.by="logFC")
head(grouped_gsva.shRAD52_STEM.shRAD52_DIFF)

## Results extraction for the fourth contrast (shNT_STEM - shNT_DIFF) sorted according to logFC
grouped_gsva.shNT_STEM.shNT_DIFF <- topTable(gsva.contrast.results, 
                      number=nrow(gsva_res),
                      coef=4,
                      sort.by="logFC")
head(grouped_gsva.shNT_STEM.shNT_DIFF)
```



---------------------- OBSOLETE -------------------------

In differentiated cells with shRAD52 (vs no shRAD52), there's a significant increase in 'HALLMARK_IL6_JAK_STAT3_SIGNALING', I should look into that. There's also differences between the two shRAD52 contrasts, depending on whether they are done on stem or diferentiated cells. 3 of the significant pathways are the same, but with shRAD52 "HALLMARK_COMPLEMENT" is present, and without it "HALLMARK_APICAL_SURFACE" is present

Update 14/05: when imputing data, the IL6_JAK_STAT3 significant increase no longer exists in differentiated cells with/without shRAD52 (p.adj = 0.133), which complicates things a little. However, the p.adj is still relatively low, and without adjusting it is significant (even though it is a bad statistical practice). When comparing expression levels of the pathway in the different conditions, a similar effect seems to exist for the stem cells (not significantly, though). We could still look into that. Also, for the STEM vs DIFF comparisons, many more significantly expressed pathways are found.

---------------------------------------------------------


Seeing which pathways are common to the two shNT vs shRAD52 comparisons, and which are unique to each comparison:
NOTE: It doesn't work because I haven't saved the significant pathways in a variable. I don't think it's so relevant anyway, can be easily looked up manually.

```{r}
gsva.shRAD52_DIFF.shNT_DIFF.significant <- gsva.shRAD52_DIFF.shNT_DIFF[]

intersect_pathways <- intersect(rownames(gsva.shNT_STEM.shNT_DIFF.significant), rownames(grouped_gsva.shRAD52_STEM.shRAD52_DIFF.significant))
unique_pathways_shNT.STEM.DIFF <- setdiff(rownames(grouped_gsva.shNT_STEM.shNT_DIFF.significant), intersect.pathways)
unique_pathways_shRAD52.STEM.DIFF <- setdiff(rownames(grouped_gsva.shRAD52_STEM.shRAD52_DIFF.significant), intersect.pathways)

```

This meta structure here created is used later for other parts like the boxplot.

```{r}
## Experimental design specification
samples <- colnames(normalized_protein)

# Create metadata
meta <- data.frame(
  sample_id = samples,
  shRNA = ifelse(grepl("^shNT", samples), "Control", "RAD52"),
  cell_type = ifelse(grepl("_DIFF_", samples), "DGC", "GSC"),
  row.names = samples
)

# Convert to factors with correct reference levels
meta$cell_type <- factor(meta$cell_type, levels = c("DGC", "GSC"))
meta$shRNA <- factor(meta$shRNA, levels = c("Control", "RAD52"))

# Making sure the metadata rows have the same order as the columns in the GSVA
meta <- meta[colnames(gsva_res), ]
```

Plotting a boxplot of a relevant pathway (could be changed later to include other pathways depending on what we find):

```{r}
df_dna_repair <- data.frame(
  score = gsva_res["HALLMARK_DNA_REPAIR", ],
  meta
)
ggplot(df_dna_repair, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_DNA_REPAIR")

df_e2f_targets <- data.frame(
  score = gsva_res["HALLMARK_E2F_TARGETS", ],
  meta
)
ggplot(df_e2f_targets, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_E2F_TARGETS")

df_jak_stat3 <- data.frame(
  score = gsva_res["HALLMARK_IL6_JAK_STAT3_SIGNALING", ],
  meta
)
ggplot(df_jak_stat3, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_IL6_JAK_STAT3_SIGNALING")

df_myc_targets <- data.frame(
  score = gsva_res["HALLMARK_MYC_TARGETS_V2", ],
  meta
)
ggplot(df_myc_targets, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_MYC_TARGETS_V2")

df_epithelial_mesenchymal <- data.frame(
  score = gsva_res["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION", ],
  meta
)
gsva_plot <- ggplot(df_epithelial_mesenchymal,
                    aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot(width = 0.5, position = position_dodge(width = 0.7)) +
  theme_minimal() +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  ggtitle("HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION")

ggsave("GSVA_EMT.svg", plot = gsva_plot, width = 6, height = 6)

df_mtor <- data.frame(
  score = gsva_res["HALLMARK_PI3K_AKT_MTOR_SIGNALING", ],
  meta
)
ggplot(df_mtor, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_PI3K_AKT_MTOR_SIGNALING")

df_oxphos <- data.frame(
  score = gsva_res["HALLMARK_OXIDATIVE_PHOSPHORYLATION", ],
  meta
)
ggplot(df_oxphos, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_OXIDATIVE_PHOSPHORYLATION")

df_uv_up <- data.frame(
  score = gsva_res["HALLMARK_UV_RESPONSE_UP", ],
  meta
)
gsva_plot <- ggplot(df_uv_up,
                    aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot(width = 0.5, position = position_dodge(width = 0.7)) +
  theme_minimal() +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  ggtitle("HALLMARK_UV_RESPONSE_UP")

ggsave("GSVA_UV_UP.svg", plot = gsva_plot, width = 6, height = 6)

df_apoptosis <- data.frame(
  score = gsva_res["HALLMARK_APOPTOSIS", ],
  meta
)
gsva_plot <- ggplot(df_apoptosis,
                    aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot(width = 0.5, position = position_dodge(width = 0.7)) +
  theme_minimal() +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  ggtitle("GSVA_APOPTOSIS.svg")

ggsave("GSVA_APOPTOSIS.svg", plot = gsva_plot, width = 6, height = 6)

df_hypoxia <- data.frame(
  score = gsva_res["HALLMARK_HYPOXIA", ],
  meta
)
ggplot(df_hypoxia, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_HYPOXIA")

df_estrogen_late <- data.frame(
  score = gsva_res["HALLMARK_ESTROGEN_RESPONSE_LATE", ],
  meta
)
ggplot(df_estrogen_late, aes(x = cell_type, y = score, fill = shRNA)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("HALLMARK_ESTROGEN_RESPONSE_LATE")
```



In the boxplot, we see reduced gene expression of the DNA Repair pathways in the RAD52 KO compared to the controls. Considering that RAD52 is involved in DNA repair, this could fit the general idea of the experiment.

Now, let's try a heatmap:

```{r}
library(pheatmap)

pathways <- rownames(na.omit(gsva_res))

pheatmap(gsva_res[pathways, ],
         annotation_col = meta[, c("cell_type", "shRNA")])
```

## Stemness

We are going to try to measure the "stemness" of the differents groups. For this, we will use different gene sets that are related to glioblastoma stem cells. These sets come from 4 different articles, each of which relate their gene set to a specific characteristic (epithelial-mesenchymal transition and radiotherapy resistance, for instance). The articles used are:

-   A CD133-related gene expression signature identifies an aggressive glioblastoma subtype with excessive mutations. <https://pmc.ncbi.nlm.nih.gov/articles/PMC3029739/>
-   A novel gene signature based on five glioblastoma stem-like cell relevant genes predicts the survival of primary glioblastoma. <https://pmc.ncbi.nlm.nih.gov/articles/PMC11813501/>
-   A Multi-Cancer Mesenchymal Transition Gene Expression Signature Is Associated with Prolonged Time to Recurrence in Glioblastoma. <https://pmc.ncbi.nlm.nih.gov/articles/PMC3321034/>
-   A Stemness-Related Gene Signature to Predict Treatment Efficacy of Radiotherapy in Glioblastoma. <https://www.sciencedirect.com/science/article/pii/S0360301622015589>

A proteomics specific paper showing their own gene signature will also be used: Okawa et al., Proteome and secretome characterisation of glioblastoma-derived neural stem cells: https://pmc.ncbi.nlm.nih.gov/articles/PMC6135235

And this one: Lottaz et al. (2010), Cancer Research

Loading the data from that paper:

```{r}
# Read the Excel sheet
okawa_data <- read_excel("Okawa_Data.xls", sheet = 1, skip = 1 )

# Clean the paper's dataset by separating the merged symbols into separate rows
okawa_expanded <- okawa_data %>%
  # Separate rows splits "AAG4;APOJ;CLU" into 3 distinct rows, copying the expression data
  separate_rows(Gene.Names, sep = ";") %>%
  # Clean up any accidental whitespace around the symbols
  mutate(Gene.Names = trimws(Gene.Names))

# Extract a clean vector of my valid gene symbols
valid_symbols <- rownames(normalized_protein)

# 3. Filter the paper's expanded dataset to ONLY keep rows that match your dataset
okawa_cleaned_and_matched <- okawa_expanded %>%
  filter(Gene.Names %in% valid_symbols)

# 4. Extract the final gene signature list for GSEA
# (Filtering for significantly upregulated proteins using the cleaned column)
okawa_stem_markers <- okawa_cleaned_and_matched %>%
  filter(`Mean 
Normalized 
mean log2 
Ratio of the 4 samples` > 1 & `Differential abundance 
Adjusted p value` < 0.05) %>%
  pull(Gene.Names) %>%
  unique()

okawa_non_stem_markers <- okawa_cleaned_and_matched %>%
  filter(`Mean 
Normalized 
mean log2 
Ratio of the 4 samples` > -1 & `Differential abundance 
Adjusted p value` < 0.05) %>%
  pull(Gene.Names) %>%
  unique()

```


### GSVA Stemness

For this, we will create the gene set lists and perform a GSVA analysis to see how each condition scores. As the two stem cell conditions are very similar (with or without RAD52 KO), this could be especially interesting in the differentiated cells with and without RAD52 KO, to see if the KO as affected the "stemness" of the cells.

```{r}
# 1. CD133+ glioblastoma stem cell signature (core subset)
GBM_CD133_STEMNESS = c(
  "AKAP4", "ARHGAP11A", "ASPM", "BARD1", "BIRC5", "BRCA1", "C12orf32", 
  "C17orf80", "C2orf48", "C4orf21", "CAPN14", "CASC5", "CCDC102A", 
  "CCDC111", "CCDC15", "CDCA2", "CDKN3", "CENPH", "CENPK", "CKAP2L", 
  "CKS2", "CTNNAL1", "DHFR", "DHX57", "DIAPH3", "DLGAP5", "DTL", 
  "DTYMK", "ECT2", "ENAH", "FANCI", "FBXO5", "GGH", "GINS2", "GMNN", 
  "H2AFZ", "HMGB2", "IFNA17", "IFNA4", "JAM2", "KIAA0101", "KIF11", 
  "KIF15", "KIF2C", "KIF4A", "KNTC1", "LIG1", "LMAN1L", "LOC91431", 
  "MAD2L1", "MCM2", "MCM3", "MELK", "MND1", "MORN2", "NACA", "NCAPH", 
  "NDC80", "NEK2", "NMU", "NUF2", "PBK", "PCNA", "POLQ", "PRIM1", 
  "PROM1", "PSG5", "PTTG1", "PTTG3", "PXMP2", "RAD51", "RANBP1", 
  "RRM2", "RTKN", "SGOL1", "SLC2A11", "SMC2", "SNRPE", "SYTL4", 
  "TIMELESS", "TM4SF1", "TMEM106C", "TOP2A", "TPX2", "TRIP13", 
  "TROAP", "TTK", "TYMS", "WDR34")

# Filter to use only the genes in our dataset
GBM_CD133_STEMNESS <- 
  GBM_CD133_STEMNESS[GBM_CD133_STEMNESS %in% rownames(normalized_protein)]

# 2. 5-gene GBM stemness signature
GBM_5GENE_STEMNESS = c(
  "ADAMTSL4",
  "CLEC5A",
  "FTSJ2",
  "IBSP",
  "SFXN4")

GBM_5GENE_STEMNESS <- 
  GBM_5GENE_STEMNESS[GBM_5GENE_STEMNESS %in% rownames(normalized_protein)]

# 3. Mesenchymal / plasticity stemness signature
GBM_MESENCHYMAL_STEMNESS = c(
  "CD44", "FN1", "VCAN", "LOX", "LOXL2",
  "COL1A1", "COL1A2", "COL5A1",
  "STEMTN", "SPARC",
  "SNAI2", "ZEB1",
  "TGFB1", "TGFBI",
  "MMP2", "MMP9")

GBM_MESENCHYMAL_STEMNESS <- 
  GBM_MESENCHYMAL_STEMNESS[GBM_MESENCHYMAL_STEMNESS %in% rownames(normalized_protein)]

# 4. 9-gene radiotherapy-associated stemness signature
GBM_RADIORESISTANCE_STEMNESS = c(
  "ASPM", "EPHA1", "LIF", "LOXL2",
  "NAP1L2", "OTP", "PITX2",
  "RDH10", "RUNX1")
  
GBM_RADIORESISTANCE_STEMNESS <- 
  GBM_RADIORESISTANCE_STEMNESS[GBM_RADIORESISTANCE_STEMNESS %in% rownames(normalized_protein)]

# Lottaz/Beier gene set
beier_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "BEIER_GLIOMA_STEM_CELL_UP") %>%
  pull(gene_symbol)

beier_non_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "BEIER_GLIOMA_STEM_CELL_DN") %>%
  pull(gene_symbol)

guenther_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "GUENTHER_GROWTH_SPHERICAL_VS_ADHERENT_UP") %>%
  pull(gene_symbol)

guenther_non_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "GUENTHER_GROWTH_SPHERICAL_VS_ADHERENT_DN") %>%
  pull(gene_symbol)

wong_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "WONG_EMBRYONIC_STEM_CELL_CORE") %>%
  pull(gene_symbol)

benporath_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "BENPORATH_ES_1") %>%
  pull(gene_symbol)

verhaak_stem_markers <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CGP") %>%
  filter(gs_name %in% "VERHAAK_GLIOBLASTOMA_PRONEURAL") %>%
  pull(gene_symbol)


gbm_stemness_sets <- list(GBM_CD133_STEMNESS,
                          # GBM_5GENE_STEMNESS,
                          # GBM_MESENCHYMAL_STEMNESS, GBM_RADIORESISTANCE_STEMNESS,
                          # okawa_stem_markers, okawa_non_stem_markers,
                          # beier_stem_markers, beier_non_stem_markers,
                          # guenther_stem_markers, guenther_non_stem_markers,
                          wong_stem_markers, benporath_stem_markers
                          # verhaak_stem_markers
                          )

names(gbm_stemness_sets) <- c(
    "Stemness_CD133_signature",
    # "Stemness5_signature",
    # "MES_signature",
    # "Radioresistance_signature",
    # "Okawa Stem Markers",
    # "Okawa Non-Stem Markers",
    # "Beier Stem Markers",
    # "Beier Non-Stem Markers",
    # "Guenther Stem Markers",
    # "Guenther Non-Stem Markers",
    "Wong Stem Markers",
    "Benporath Stem Markers"
    # "Verhaak Stem Markers"
    )
```

We can check how many of the genes in the gene sets are actually in our proteomic data:

```{r}
lapply(gbm_stemness_sets, function(gs) {
  sum(gs %in% rownames(normalized_protein))
})
```

We see that from the second gene set, only 1 out of 5 genes are actually present in the data. From the fourth gene set, 3 out of 9 are present. This may make the results from these datasets unstable. Perhaps we will have to focus on the other two.

```{r}
param_stemness <- gsvaParam(
  exprData = expr_mat,
  geneSets = gbm_stemness_sets,
  kcdf = "Gaussian"  # Using ssGSEA, which is in theory more robust for small gene sets
)

gsva_stemness_res <- gsva(param_stemness)

## Linear model fit to the experimental design, mean expression calculation
gsva.stem.linear.fit <- lmFit(gsva_stemness_res, experimental.design)

## Contrast specification treatment-control
gsva.stem.contrast.matrix <- makeContrasts(shRAD52_DIFF-shNT_DIFF,shRAD52_STEM-shNT_STEM,
                                      shRAD52_STEM-shRAD52_DIFF,shNT_STEM-shNT_DIFF,
                                      
                                 levels=c("shNT_DIFF","shNT_STEM","shRAD52_DIFF","shRAD52_STEM"))

gsva.stem.contrast.linear.fit <- contrasts.fit(gsva.stem.linear.fit, gsva.stem.contrast.matrix)
gsva.stem.contrast.results <- eBayes(gsva.stem.contrast.linear.fit)

## Results extraction for the first contrast (shRAD52_DIFF - shNT_DIFF) sorted according to logFC
gsva.stem.shRAD52_DIFF.shNT_DIFF <- topTable(gsva.stem.contrast.results, 
                      number=nrow(gsva_stemness_res),
                      coef=1,
                      sort.by="logFC")
gsva.stem.shRAD52_DIFF.shNT_DIFF

## Results extraction for the second contrast (shRAD52_STEM - shNT_STEM) sorted according to logFC
gsva.stem.shRAD52_STEM.shNT_STEM <- topTable(gsva.stem.contrast.results, 
                      number=nrow(gsva_res),
                      coef=2,
                      sort.by="logFC")
gsva.stem.shRAD52_STEM.shNT_STEM

## Results extraction for the fourth contrast (shNT_STEM - shNT_DIFF) sorted according to logFC
gsva.stem.shNT_STEM.shNT_DIFF <- topTable(gsva.stem.contrast.results, 
                      number=nrow(gsva_res),
                      coef=4,
                      sort.by="logFC")
gsva.stem.shNT_STEM.shNT_DIFF
```

So between the stem cells and the differentiated ones, there are expected significant changes in the gene sets (except dataset 2, in which only 1 gene was present in our study and therefore isn't reliable). However, between the RAD52 KO and non KO of the differentiated cells, no significant differences in stemness are found. There could be a diminished stemness after shRAD52 treatment according to the radiotherapy geneset, the 4th one (logFC = -0.52684738, p.val = 0.04264624, adj.p.val = 0.1705850), but it is not especially reliable with only 3 genes being used for that comparison.

### GSEA Stemness

Calculating stemness, but with GSEA methods instead of GSVA

```{r}
# RAD52 DIFF
ranks <- shRAD52_DIFF.shNT_DIFF$t
names(ranks) <- rownames(shRAD52_DIFF.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

stemness.fgseaRes.shRAD52_DIFF.shNT_DIFF <- fgsea(
  pathways = gbm_stemness_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

stemness.fgseaRes.shRAD52_DIFF.shNT_DIFF

# RAD52 STEM
ranks <- shRAD52_STEM.shNT_STEM$t
names(ranks) <- rownames(shRAD52_STEM.shNT_STEM)

ranks <- sort(ranks, decreasing = TRUE)

stemness.fgseaRes.shRAD52_STEM.shNT_STEM <- fgsea(
  pathways = gbm_stemness_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

stemness.fgseaRes.shRAD52_STEM.shNT_STEM

# STEM - DIFF

ranks <- shNT_STEM.shNT_DIFF$t
names(ranks) <- rownames(shNT_STEM.shNT_DIFF)

ranks <- sort(ranks, decreasing = TRUE)

stemness.fgseaRes.shNT_STEM.shNT_DIFF <- fgsea(
  pathways = gbm_stemness_sets,
  stats = ranks,
  minSize = 5,
  maxSize = 500)

stemness.fgseaRes.shNT_STEM.shNT_DIFF
```

```{r}
interactors <- read.csv(file = "BIOGRID-GENE-111830-5.0.257.tab3.txt", sep = '\t')
```

## Transcription Factor Enrichment

### TF Enrichment Calculation

We will try infer the activated Transcription Factors (TF) from the protein expression dataset using Dorothea and decoupleR.

```{r}
# Using the t-statistic as a measure of change
stat_vec <- shNT_STEM.shNT_DIFF$t
names(stat_vec) <- rownames(shNT_STEM.shNT_DIFF)

# Loading the TF data from dorothea
data(dorothea_hs, package = "dorothea")

# Keeping only higher-confidence interactions (dorothea classifies the
# confidence of the interactions in 5 categories, A-E):
regulons <- dorothea_hs %>%
  filter(confidence %in% c("A","B"))

# Creating an input matrix for decoupleR
mat <- matrix(
  stat_vec,
  ncol = 1,
  dimnames = list(names(stat_vec), "STEM_vs_DIFF"))

# We will run several models and see the consensus features 

TF_ulm_STEM_DIFF <- run_ulm(
  mat = mat,
  network = regulons,
  .source = "tf",
  .target = "target",
  .mor = "mor")

TF_mlm_STEM_DIFF <- run_mlm(
  mat = mat,
  network = regulons,
  .source = "tf",
  .target = "target",
  .mor = "mor")

# Now, to select only those which are significant in both methods, for higher
# confidence
significant_TF_mlm_STEM_DIFF <- TF_mlm_STEM_DIFF[TF_mlm_STEM_DIFF$p_value<0.05,]
significant_TF_ulm_STEM_DIFF <- TF_ulm_STEM_DIFF[TF_ulm_STEM_DIFF$p_value<0.05,]

intersection_TF_STEM_DIFF <- intersect(significant_TF_mlm_STEM_DIFF$source,significant_TF_ulm_STEM_DIFF$source)
```

For the RAD52 vs NT comparison in STEM cells:

```{r}
# Using the t-statistic as a measure of change
stat_vec <- shRAD52_STEM.shNT_STEM$t
names(stat_vec) <- rownames(shRAD52_STEM.shNT_STEM)

# Loading the TF data from dorothea
data(dorothea_hs, package = "dorothea")

# Keeping only higher-confidence interactions (dorothea classifies the
# confidence of the interactions in 5 categories, A-E):
regulons <- dorothea_hs %>%
  filter(confidence %in% c("A","B"))

# Creating an input matrix for decoupleR
mat <- matrix(
  stat_vec,
  ncol = 1,
  dimnames = list(names(stat_vec), "RAD52_vs_NT"))

# We will run several models and see the consensus features 

TF_ulm_RAD52_STEM <- run_ulm(
  mat = mat,
  network = regulons,
  .source = "tf",
  .target = "target",
  .mor = "mor")

TF_mlm_RAD52_STEM <- run_mlm(
  mat = mat,
  network = regulons,
  .source = "tf",
  .target = "target",
  .mor = "mor")

# Now, to select only those which are significant in both methods, for higher
# confidence
significant_TF_mlm_RAD52_STEM <- TF_mlm_RAD52_STEM[TF_mlm_RAD52_STEM$p_value<0.05,]
significant_TF_ulm_RAD52_STEM <- TF_ulm_RAD52_STEM[TF_ulm_RAD52_STEM$p_value<0.05,]

intersection_TF_RAD52_STEM <- intersect(significant_TF_mlm_RAD52_STEM$source,significant_TF_ulm_RAD52_STEM$source)
```

And for the RAD52 vs NT comparison in DIFF cells:

```{r}
# Using the t-statistic as a measure of change
stat_vec <- shRAD52_DIFF.shNT_DIFF$t
names(stat_vec) <- rownames(shRAD52_DIFF.shNT_DIFF)

# Loading the TF data from dorothea
data(dorothea_hs, package = "dorothea")

# Keeping only higher-confidence interactions (dorothea classifies the
# confidence of the interactions in 5 categories, A-E):
regulons <- dorothea_hs %>%
  filter(confidence %in% c("A","B","C"))  # Including category C because of lack
                                          # of results using only A and B

# Creating an input matrix for decoupleR
mat <- matrix(
  stat_vec,
  ncol = 1,
  dimnames = list(names(stat_vec), "RAD52_vs_NT"))

# We will run several models and see the consensus features 

TF_ulm_RAD52_DIFF <- run_ulm(
  mat = mat,
  network = regulons,
  .source = "tf",
  .target = "target",
  .mor = "mor")

TF_mlm_RAD52_DIFF <- run_mlm(
  mat = mat,
  network = regulons,
  .source = "tf",
  .target = "target",
  .mor = "mor")

# Now, to select only those which are significant in both methods, for higher
# confidence
significant_TF_mlm_RAD52_DIFF <- TF_mlm_RAD52_DIFF[TF_mlm_RAD52_DIFF$p_value<0.05,]
significant_TF_ulm_RAD52_DIFF <- TF_ulm_RAD52_DIFF[TF_ulm_RAD52_DIFF$p_value<0.05,]
intersection_TF_RAD52_DIFF <- intersect(significant_TF_mlm_RAD52_DIFF$source,significant_TF_ulm_RAD52_DIFF$source)
```

Dotplot, barplot, volcanoplot

```{r}
thap11_targets <- regulons %>%
  dplyr::filter(tf == "THAP11")

## Visualization1

target_count <- regulons %>%
  filter(target %in% rownames(mat)) %>%
  count(tf, name = "n_targets")

plot_df <- TF_mlm_RAD52_DIFF %>%
  left_join(target_count,
            by = c("source" = "tf")) %>%
  mutate(
    neglog10p = -log10(p_value),
    Significant = p_value < 0.05) %>%
  filter(p_value < 0.05)

plot_df$Label <- ifelse(
    plot_df$p_value < 0.02 &
    abs(plot_df$score) > 2,
    plot_df$source,
    ""
)

ggplot(plot_df,
       aes(score,
           reorder(source, score))) +

    geom_point(aes(size = n_targets,
                   colour = neglog10p),
               alpha = 0.9) +

    geom_text_repel(aes(label = Label),
                    size = 3.5,
                    max.overlaps = Inf) +

    scale_colour_gradient(
        low = "grey80",
        high = "#D73027",
        name = expression(-log[10](italic(P)))
    ) +

    scale_size(
        range = c(2,10),
        name = "Detected\ntargets"
    ) +

    geom_vline(xintercept = 0,
               linetype = 2,
               colour = "grey60") +

    labs(x = "TF activity score (ULM)",
         y = NULL) +

    theme_classic(base_size = 14) +

    theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "right"
    )

# Barplot
ggplot(plot_df,
       aes(x = score,
           y = reorder(source, score))) +
    geom_col() +
    theme_classic()
```

Consensus MLM+ULM plot (RAD52_DIFF)

```{r}
regulons <- dorothea_hs %>%
  filter(confidence %in% c("A","B","C"))

target_count <- regulons %>%
  filter(target %in% rownames(mat)) %>%
  count(tf, name = "n_targets")

consensus_df <- TF_ulm_RAD52_DIFF %>%
  dplyr::select(source,
         ulm_score = score,
         ulm_p = p_value) %>%
  inner_join(
    TF_mlm_RAD52_DIFF %>%
      dplyr::select(source,
             mlm_score = score,
             mlm_p = p_value),
    by = "source"
  ) %>%
  left_join(
    target_count,
    by = c("source" = "tf")
  )

consensus_df <- consensus_df %>%
  mutate(
    consensus_p = pmax(ulm_p, mlm_p),      # the less significant p-value
    neglog10p = -log10(consensus_p),

    significant =
      ulm_p < 0.05 &
      mlm_p < 0.05,

    label = ifelse(significant, source, "")
  )

consensus_df <- consensus_df %>%
  mutate(mean_score = (ulm_score + mlm_score)/2)

consensus_rad52diff_plot <- ggplot(consensus_df,
       aes(x = ulm_score,
           y = mlm_score,
           colour = mean_score)) +
  
    # --- quadrant shading ---
  annotate("rect",
           xmin = -Inf, xmax = 0,
           ymin = -Inf, ymax = 0,
           fill = "#3B6FB6", alpha = 0.08) +

  annotate("rect",
           xmin = 0, xmax = Inf,
           ymin = 0, ymax = Inf,
           fill = "#C0392B", alpha = 0.08) +

  ################
  
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "grey80") +

  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey80") +

  geom_abline(intercept = 0,
              slope = 1,
              colour = "grey60",
              linewidth = 0.5) +

  geom_point(
    aes(size = n_targets,
        colour = neglog10p),
    alpha = 0.85
  ) +

  geom_text_repel(
  aes(label = label),
  fontface = "bold",
  size = 4,
  box.padding = 0.5,
  point.padding = 0.4,
  segment.color = "grey60",
  segment.size = 0.4,
  max.overlaps = Inf
  ) +

  scale_colour_gradient2(
    low = "blue",
    mid = "grey92",
    high = "red",
    midpoint = 0
  ) +

  scale_size(
    range = c(2,10),
    name = "Detected\ntargets"
  ) +

  labs(
    x = "ULM activity score",
    y = "MLM activity score"
  ) +

  #coord_equal() +
  
  coord_cartesian(xlim = c(-3.2, 3.4), ylim = c(-2.9, 3.4)) +
  
  geom_point(
    data = subset(consensus_df, significant),
    shape = 21,
    stroke = 1.2,
    colour = "black",
    fill = NA,
    aes(size = n_targets)
  ) +

  theme_classic(base_size = 14)

ggsave("TF_RAD52DIFF.svg", plot = consensus_rad52diff_plot, width = 8, height = 6)
```

Consensus MLM+ULM plot (RAD52_STEM)

```{r}
regulons <- dorothea_hs %>%
  filter(confidence %in% c("A","B"))

target_count <- regulons %>%
  filter(target %in% rownames(mat)) %>%
  count(tf, name = "n_targets")

consensus_df <- TF_ulm_RAD52_STEM %>%
  dplyr::select(source,
         ulm_score = score,
         ulm_p = p_value) %>%
  inner_join(
    TF_mlm_RAD52_STEM %>%
      dplyr::select(source,
             mlm_score = score,
             mlm_p = p_value),
    by = "source"
  ) %>%
  left_join(
    target_count,
    by = c("source" = "tf")
  )

consensus_df <- consensus_df %>%
  mutate(
    consensus_p = pmax(ulm_p, mlm_p),      # the less significant p-value
    neglog10p = -log10(consensus_p),

    significant =
      ulm_p < 0.05 &
      mlm_p < 0.05,

    label = ifelse(significant, source, "")
  )

consensus_df <- consensus_df %>%
  mutate(mean_score = (ulm_score + mlm_score)/2)

consensus_rad52stem_plot <- ggplot(consensus_df,
       aes(x = ulm_score,
           y = mlm_score,
           colour = mean_score)) +
  
    # --- quadrant shading ---
  annotate("rect",
           xmin = -Inf, xmax = 0,
           ymin = -Inf, ymax = 0,
           fill = "#3B6FB6", alpha = 0.08) +

  annotate("rect",
           xmin = 0, xmax = Inf,
           ymin = 0, ymax = Inf,
           fill = "#C0392B", alpha = 0.08) +

  ################
  
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "grey80") +

  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey80") +

  geom_abline(intercept = 0,
              slope = 1,
              colour = "grey60",
              linewidth = 0.5) +

  geom_point(
    aes(size = n_targets,
        colour = neglog10p),
    alpha = 0.85
  ) +

  geom_text_repel(
  aes(label = label),
  fontface = "bold",
  size = 4,
  box.padding = 0.5,
  point.padding = 0.4,
  segment.color = "grey60",
  segment.size = 0.4,
  max.overlaps = Inf
  ) +

  scale_colour_gradient2(
    low = "blue",
    mid = "grey92",
    high = "red",
    midpoint = 0
  ) +

  scale_size(
    range = c(2,10),
    name = "Detected\ntargets"
  ) +

  labs(
    x = "ULM activity score",
    y = "MLM activity score"
  ) +

  coord_equal() +
  
  #coord_cartesian(xlim = c(-3.2, 3.4), ylim = c(-2.9, 3.4)) +
  
  geom_point(
    data = subset(consensus_df, significant),
    shape = 21,
    stroke = 1.2,
    colour = "black",
    fill = NA,
    aes(size = n_targets)
  ) +

  theme_classic(base_size = 14)

ggsave("TF_RAD52STEM.svg", plot = consensus_rad52stem_plot, width = 8, height = 6)
```

The following code checks how many targets of each TF is covered in our protein dataset.

```{r}
covered_targets <- regulons %>%
  filter(target %in% rownames(mat)) %>%
  count(tf)
```

### Visualization

```{r}

```

Heatmap (doesn't work in this case because it needs at least 2 different objects to cluster, but the code might be useful):

```{r}
n_tfs <- 15

# Transform to wide matrix
sample_acts_mat <- res_ulm_STEM_DIFF %>%
                   tidyr::pivot_wider(id_cols = 'condition', 
                                      names_from = 'source',
                                      values_from = 'score') %>%
                   tibble::column_to_rownames('condition') %>%
                   as.matrix()

# Get top tfs with more variable means across clusters
tfs <- res_ulm_STEM_DIFF %>%
       dplyr::group_by(source) %>%
       dplyr::summarise(std = stats::sd(score)) %>%
       dplyr::arrange(-abs(std)) %>%
       head(n_tfs) %>%
       dplyr::pull(source)

sample_acts_mat <- sample_acts_mat[,tfs]

# Scale per sample
sample_acts_mat <- scale(sample_acts_mat)

# Choose color palette
colors <- rev(RColorBrewer::brewer.pal(n = 11, name = "RdBu"))
colors.use <- grDevices::colorRampPalette(colors = colors)(100)

my_breaks <- c(seq(-2, 0, length.out = ceiling(100 / 2) + 1),
               seq(0.05, 2, length.out = floor(100 / 2)))

# Plot
pheatmap::pheatmap(mat = sample_acts_mat,
                   color = colors.use,
                   border_color = "white",
                   breaks = my_breaks,
                   cellwidth = 15,
                   cellheight = 15,
                   treeheight_row = 20,
                   treeheight_col = 20)
```

Infer TF activities from the t-values of the DEGs between STEM and DIFF:

```{r}
# Filter top TFs in both signs
f_contrast_acts <- res_ulm %>%
                   dplyr::mutate(rnk = NA)

msk <- f_contrast_acts$score > 0

f_contrast_acts[msk, 'rnk'] <- rank(-f_contrast_acts[msk, 'score'])
f_contrast_acts[!msk, 'rnk'] <- rank(-abs(f_contrast_acts[!msk, 'score']))

tfs <- f_contrast_acts %>%
       dplyr::arrange(rnk) %>%
       head(n_tfs) %>%
       dplyr::pull(source)

f_contrast_acts <- f_contrast_acts %>%
                   filter(source %in% tfs)

colors <- rev(RColorBrewer::brewer.pal(n = 11, name = "RdBu")[c(2, 10)])

p <- ggplot2::ggplot(data = f_contrast_acts, 
                     mapping = ggplot2::aes(x = stats::reorder(source, score), 
                                            y = score)) + 
     ggplot2::geom_bar(mapping = ggplot2::aes(fill = score),
                       color = "black",
                       stat = "identity") +
     ggplot2::scale_fill_gradient2(low = colors[1], 
                                   mid = "whitesmoke", 
                                   high = colors[2], 
                                   midpoint = 0) + 
     ggplot2::theme_minimal() +
     ggplot2::theme(axis.title = element_text(face = "bold", size = 12),
              axis.text.x = ggplot2::element_text(angle = 45, 
                                                  hjust = 1, 
                                                  size = 10, 
                                                  face = "bold"),
              axis.text.y = ggplot2::element_text(size = 10, 
                                                  face = "bold"),
              panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank()) +
     ggplot2::xlab("TFs")

p

```

```{r}
ggplot(TF_ulm_RAD52_DIFF,
       aes(x = score,
           y = reorder(source, score))) +
    geom_col() +
    theme_classic()
```

```{r}
covered_targets <- regulons %>%
  filter(target %in% rownames(mat)) %>%
  count(tf)
```


## Test: common activated and repressed

```{r}
common_activated_DIFF_STEM <- intersect(shRAD52_DIFF.shNT_DIFF.activated.genes, shRAD52_STEM.shNT_STEM.activated.genes)
common_repressed_DIFF_STEM <- intersect(shRAD52_DIFF.shNT_DIFF.repressed.genes, shRAD52_STEM.shNT_STEM.repressed.genes)
```

## Leading Edge genes (GSEA)

Checking Leading Edge genes in the topTable df:

```{r}
# Function to extract the leading-edge genes of one geneset/pathway in a
# fgsea contrast. It then creates a table with a description for each gene
# in the leading-edge

leading_edge_function <- function(fgsea_result, pathway, contrast, filename){
  # See at which row the pathway is indexed and extract the leading-edge genes
  geneset_index <- which(fgsea_result[,1] == pathway)
  leading_edge <- fgsea_result[[geneset_index,8]]
  
  # Create the annotation (description) of each gene
  annot <- queryMany(
    leading_edge,
    scopes = "symbol",
    fields = c("name", "summary"),
    species = "human")
  
  # Save the annotation as a .tsv table
  annot <- data.frame(annot$query, annot$name, annot$summary)
  write.table(x = annot, file = filename, quote = F,
            row.names = F, sep = "\t")
  
  # Returns a table with the TopTable results of the leading-edge genes, so that
  # we can easily check the p-values of those genes
  return(contrast[leading_edge,])
}

leading_edge_function(fgsea_result = fgseaRes.shNT_STEM.shNT_DIFF,
                      pathway = "HALLMARK_DNA_REPAIR",
                      contrast = shNT_STEM.shNT_DIFF,
                      filename = "leading_edge_STEMDIFF_DNA_REPAIR.tsv")
```

RAD52 vs NT, DIFF cells

```{r}
leading_edge_function(fgsea_result = fgseaRes.shRAD52_DIFF.shNT_DIFF,
                      pathway = "HALLMARK_HYPOXIA",
                      contrast = shRAD52_DIFF.shNT_DIFF,
                      filename = "leading_edge_RAD52DIFF_HYPOXIA.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_DIFF.shNT_DIFF,
                      pathway = "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
                      contrast = shRAD52_DIFF.shNT_DIFF,
                      filename = "leading_edge_RAD52DIFF_EMT.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_DIFF.shNT_DIFF,
                      pathway = "HALLMARK_ANGIOGENESIS",
                      contrast = shRAD52_DIFF.shNT_DIFF,
                      filename = "leading_edge_RAD52DIFF_ANGIOGENESIS.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_DIFF.shNT_DIFF,
                      pathway = "HALLMARK_UV_RESPONSE_UP",
                      contrast = shRAD52_DIFF.shNT_DIFF,
                      filename = "leading_edge_RAD52DIFF_UV_UP.tsv")

## fry() identifies some pathways that fgesa() doesn't. Let's look at the genes
## of those fry()-exclusive pathways

p53_genes_RAD52DIFF <- na.omit(shRAD52_DIFF.shNT_DIFF[gene_sets[["HALLMARK_P53_PATHWAY"]],])
p53_genes_RAD52DIFF[order(p53_genes_RAD52DIFF$P.Value),]

mTOR_genes_RAD52DIFF <- na.omit(shRAD52_DIFF.shNT_DIFF[gene_sets[["HALLMARK_PI3K_AKT_MTOR_SIGNALING"]],])
mTOR_genes_RAD52DIFF[order(mTOR_genes_RAD52DIFF$P.Value),]


```

RAD52 vs NT, STEM cells

```{r}
leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_EMT.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_ESTROGEN_RESPONSE_LATE",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_ESTROGEN.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_COAGULATION",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_COAGULATION.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_APOPTOSIS",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_APOPTOSIS.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_COMPLEMENT",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_COMPLEMENT.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_INFLAMMATORY_RESPONSE",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_INFLAMMATORY.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_PI3K_AKT_MTOR_SIGNALING",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_MTOR.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_UV_RESPONSE_UP",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_UV_UP.tsv")

leading_edge_function(fgsea_result = fgseaRes.shRAD52_STEM.shNT_STEM,
                      pathway = "HALLMARK_KRAS_SIGNALING_UP",
                      contrast = shRAD52_STEM.shNT_STEM,
                      filename = "leading_edge_RAD52STEM_KRAS.tsv")


UV_UP_genes_RAD52STEM <- na.omit(shRAD52_STEM.shNT_STEM[gene_sets[["HALLMARK_UV_RESPONSE_UP"]],])
UV_UP_genes_RAD52STEM[order(UV_UP_genes_RAD52STEM$P.Value),]

DNAREPAIR_genes_RAD52STEM <- na.omit(shRAD52_STEM.shNT_STEM[gene_sets[["HALLMARK_DNA_REPAIR"]],])
DNAREPAIR_genes_RAD52STEM[order(DNAREPAIR_genes_RAD52STEM$P.Value),]

```

### Heatmaps

Heatmap of protein expression for the GSEA results

```{r}
heatmap_gsea <- function(pathway, protein_data, samples, image_name){
  # Careful: requieres gene_sets to be defined before running the function.
  # Gets the list of genes from the pathway and extracts the expression values
  # from the protein data df
  geneset_prots <- gene_sets[[pathway]]
  # The input argument "samples" has to be a vector specifying the indexes of
  # the samples we are mapping in the protein data df
  protein_matrix <- na.omit(protein_data[geneset_prots, samples])
  
  # Draws and saves the heatmap as svg and jpg
  heatmap <- pheatmap::pheatmap(mat = protein_matrix,
                   scale = 'row', # To scale each protein individually, in case
                                  # one has much higher expression than the rest
                   color = colors.use,
                   border_color = "white")
  
  ggplot2::ggsave(paste0(image_name,".svg"),
                heatmap$gtable,
                width = 8,
                height = 20)

  ggplot2::ggsave(paste0(image_name,".jpg"),
                heatmap$gtable,
                width = 8,
                height = 20,
                dpi = 300)
}

# The same function, but only representing the leading edge genes of the pathway

heatmap_gsea_leading <- function(pathway, protein_data, samples, image_name, fgsea_result){
  # Careful: requieres gene_sets to be defined before running the function.
  # Gets the list of genes from the pathway and extracts the expression values
  # from the protein data df
  geneset_index <- which(fgsea_result[,1] == pathway)
  leading_edge <- fgsea_result[[geneset_index,8]]

  # The input argument "samples" has to be a vector specifying the indexes of
  # the samples we are mapping in the protein data df
  protein_matrix <- na.omit(protein_data[leading_edge, samples])
  
  # Draws and saves the heatmap as svg and jpg
  heatmap <- pheatmap::pheatmap(mat = protein_matrix,
                   scale = 'row', # To scale each protein individually, in case
                                  # one has much higher expression than the rest
                   color = colors.use,
                   border_color = "white")
  
  ggplot2::ggsave(paste0(image_name,".svg"),
                heatmap$gtable,
                width = 8,
                height = 20)

  ggplot2::ggsave(paste0(image_name,".jpg"),
                heatmap$gtable,
                width = 8,
                height = 20,
                dpi = 300)
}

heatmap_gsea(pathway = "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_EMT")

heatmap_gsea(pathway = "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_OXPHOS")

heatmap_gsea(pathway = "HALLMARK_DNA_REPAIR",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_DNAREPAIR")

heatmap_gsea_leading(pathway = "HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_EMT_leading",
             fgsea_result = fgseaRes.shNT_STEM.shNT_DIFF)

heatmap_gsea_leading(pathway = "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_OXPHOS_leading",
             fgsea_result = fgseaRes.shNT_STEM.shNT_DIFF)

heatmap_gsea_leading(pathway = "HALLMARK_DNA_REPAIR",
             protein_data = normalized_protein,
             samples = c(1:8),
             image_name = "heatmap_STEMDIFF_DNAREPAIR_leading",
             fgsea_result = fgseaRes.shNT_STEM.shNT_DIFF)
```

Heatmap of protein expression for the OXPHOS proteins

```{r}
# OXPHOS_prots <- pathway_contrast
# OXPHOS_prots_pval <- OXPHOS_prots$P.Value

# OXPHOS_prots <- gene_sets[["HALLMARK_OXIDATIVE_PHOSPHORYLATION"]]

OXPHOS_prots <- fgseaRes.shNT_STEM.shNT_DIFF[[36,8]]
OXPHOS_prots <- shNT_STEM.shNT_DIFF[OXPHOS_prots,]
# OXPHOS_prots_padj <- OXPHOS_prots$adj.P.Val
# OXPHOS_prots <- OXPHOS_prots[OXPHOS_prots_padj < 0.05,]
OXPHOS_prots <- rownames(OXPHOS_prots)

OXPHOS_matrix <- na.omit(normalized_protein[OXPHOS_prots,c(1:8)])

svg("heatmap_OXPHOS.svg", width = 8, height = 16)
# jpeg("heatmap_OXPHOS.jpg", width = 1600, height = 3200, res = 300)
pheatmap::pheatmap(mat = OXPHOS_matrix,
                   scale = 'row', # To scale each protein individually, in case
                                  # one has much higher expression than the rest
                   color = colors.use,
                   border_color = "white",
                   # cellwidth = 15,
                   # cellheight = 15,
                   # treeheight_row = 20,
                   # treeheight_col = 20
                   )
dev.off()
```

```{r}
thap11_prots <- thap11_targets$target
thap11_matrix <- na.omit(normalized_prot_DIFF[thap11_prots,])

pheatmap::pheatmap(mat = thap11_matrix,
                   scale = 'row', # To scale each protein individually, in case
                                  # one has much higher expression than the rest
                   color = colors.use,
                   border_color = "white",
                   # cellwidth = 15,
                   # cellheight = 15,
                   # treeheight_row = 20,
                   # treeheight_col = 20
                   )
```